---
title: "R codes for the NutNet-Stability project"
subtitle: "Step3: Analysis on single stability components"
author: "Qiang Yang"
date: "`r Sys.Date()`"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{}
- \fancyfoot[CO,CE]{R codes for the NutNet-stability project}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage


In this section, I conduct regression analysis on single stability components to see their response to nutrient enrichment and their dependence on time length.

# Section 1: set up working environment 

```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, fig.width=7, fig.height=4}
rm(list = ls())
gc()
# necessary R packages
Packages <- c(
  "tidyverse", "magrittr", "grid", "gridExtra",
  "scales", "broom", "nlme", "sjPlot", "patchwork", "lme4", "lmerTest"
)
pacman::p_load(char = Packages)
# functions with conflicting status (presence in different packages used here)
select <- dplyr::select
summarise <- dplyr::summarise

# working path
main.path <- getwd()
data.path <- paste0(main.path, "/mydata")
figure.path <- paste0(main.path, "/myfigure")
set.seed(1234)
```

# Section 2 prepare data

```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, fig.width=7, fig.height=4}
load(paste0(data.path, "/df.tidy.stabilities.Rdata"))
load(paste0(data.path, "/df.tidy.Rdata"))
df.full <- df.tidy.stabilities %>%
  left_join(df.tidy %>% select(-contains("data")))
rm(df.tidy, df.tidy.stabilities)


df.full %<>%
  select(site_code:end_year, matches("_added_proportion|change|var"))
sta.names <- colnames(df.full) %>% str_subset("change|var")

# scale the stability components here
# some of the columns include Inf and NA, so make a new scale function
new.scale <- function(x) {
  x[is.infinite(x)] <- NA
  y <- scale(x)
}
df.full <- df.full %>%
  mutate(
    func_var = new.scale(func_var),
    func_var_det = new.scale(func_var_det),
    comp_var = new.scale(comp_var),
    func_change = new.scale(func_change),
    comp_change = new.scale(comp_change)
  )

# data for the treatments N, P, and K
df.meta.N.P.K <- data.frame(trt = c("N", "P", "K")) %>%
  mutate(pert.name = paste0(trt, "_added_proportion")) %>%
  expand_grid(sta.names)

df.N.P.K.nested <- df.meta.N.P.K %>%
  mutate(
    data = pmap(list(trt, pert.name, sta.names), function(x1, x2, x3) {
      output <- df.full %>%
        filter(trt == x1) %>%
        select(one_of(c("site_code", "block", "plot", "end_year", x2, x3))) %>%
        set_colnames(c("site_code", "block", "plot", "end_year", "perturbation_intensity", "stability"))
    })
  )

# note that there will be missing data in the stability column
# this is due to the availability of cover data or biomass data for some plots
# remove the NA values
df.N.P.K.nested <- df.N.P.K.nested %>%
  mutate(data = map(data, na.omit))
# I also noted that there are infinite values in the stability column
#  remove them
df.N.P.K.nested <- df.N.P.K.nested %>%
  mutate(data = map(data, function(df1) df1 %>% filter(!is.infinite(stability))))
```

# Section 3 - sites as random intercept

I run two sets of model. The first: using sites as random intercept; the second: using sites as both random intercept and random slopes.

**3.1 perform the regression**
```{r, warning=FALSE, message=FALSE,  highlight=T, fig.width=7, fig.height=4, eval=TRUE}
df.models.N.P.K <- df.N.P.K.nested %>%
  mutate(mod.out = map(data, function(dat) {
    # log transform perturbation intensity
    dat %<>% mutate(log.pert = log(perturbation_intensity))
    dat %<>% na.omit()
    mod.dat <- dat
    mod.dat %<>% mutate(across(c(end_year, log.pert), scale))
    formula_ <- stability ~ log.pert + end_year + log.pert:end_year + (1 | site_code)
    mod <- lmer(formula = formula_, data = mod.dat)
    return(mod)
  }))
```
There is no convergence problem for this set of models.

**3.2 organize model coefficient estimates and p values**

```{r, warning=FALSE, message=FALSE,  highlight=T, fig.width=7, fig.height=4, eval=TRUE}
df.models.N.P.K.coeffficients <- df.models.N.P.K %>%
  mutate(mod.coefficient = map(mod.out, function(mod) {
    # coefficient estimate from the summary function
    output <- summary(mod)$coefficients
    coeff <- output %>% as.data.frame()
    coeff %<>% mutate(predictor = rownames(output))
    coeff %<>% select(predictor, Estimate)
    # confidence interval from lme4::confint.merMod()
    # profile confidence intervals
    cfs <- confint.merMod(mod)
    cfs %<>% as.data.frame()
    cfs %<>% mutate(predictor = rownames(cfs))
    cfs %<>% rename(lcf = `2.5 %`, ucf = `97.5 %`)
    cfs %<>% filter(!str_detect(predictor, ".sig"))
    # P value from anova (lmerTest borrow from pkgtest)
    df.p.value <- anova(mod) %>% as.data.frame()
    df.p.value %<>% select(NumDF:`Pr(>F)`)
    df.p.value %<>% mutate(predictor = rownames(df.p.value))

    # add significance sign
    coeff <- coeff %>%
      left_join(cfs) %>%
      left_join(df.p.value)
    coeff %<>% filter(predictor != "(Intercept)")
    coeff %<>%
      mutate(sig.sign = map2_chr(Estimate, `Pr(>F)`, function(x, p) {
        if (x < 0 & p < 0.05) {
          out <- "negative"
        } else if (x > 0 & p < 0.05) {
          out <- "positive"
        } else {
          out <- "non-significant"
        }
        return(out)
      }))
    return(coeff)
  }))


df.models.N.P.K.coeffficients %<>%
  mutate(model.ID = 1:nrow(df.models.N.P.K.coeffficients)) %>%
  select(-pert.name, -data, -mod.out) %>%
  select(model.ID, everything()) %>%
  unnest(cols = mod.coefficient)
```

**3.3 visualize model coefficient estimates**

```{r, warning=FALSE, message=FALSE, highlight=T, fig.width=6, fig.height=4, eval=TRUE}
df.coefficients <- df.models.N.P.K.coeffficients %>%
  mutate(predictor = str_replace_all(predictor, "end_year", "years")) %>%
  mutate(predictor = str_replace_all(predictor, "log.pert", "perturbation"))
stabilities <- df.coefficients %>%
  select(sta.names) %>%
  distinct() %>%
  pull()
df.coefficients %<>%
  mutate(sta.names = as.factor(sta.names)) %>%
  mutate(sta.names = fct_relevel(sta.names, stabilities)) %>%
  mutate(trt = as.factor(trt)) %>%
  mutate(trt = fct_relevel(trt, c("N", "P", "K"))) %>%
  mutate(predictor = as.factor(predictor)) %>%
  mutate(predictor = fct_relevel(predictor, c("years", "perturbation", "perturbation:years")))

p1 <- df.coefficients %>%
  # filter(str_detect(sta.names, "functional")) %>%
  ggplot(aes(x = predictor, y = Estimate, color = sig.sign)) +
  geom_point() +
  geom_errorbar(aes(ymin = lcf, ymax = ucf), width = .2) +
  facet_grid(trt ~ sta.names) +
  coord_flip() +
  geom_hline(yintercept = 0, col = "grey", linetype = "dashed", size = 0.3) +
  scale_color_manual(values = c("blue", "grey", "red")) +
  theme_linedraw() +
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(color = "black", size = 8),
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 9, color = "black", angle = 90),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_blank(),
    plot.title = element_text(face = "bold", color = "blue"),
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  xlab("Predictors") +
  ylab("The Coefficient Estimate")

trt.labs <- c("+N", "+P", "+K")
names(trt.labs) <- c("N", "P", "K")

# Create the plot
(p1 <- p1 + facet_grid(
  trt ~ sta.names,
  labeller = labeller(trt = trt.labs)
))

# make a table for the coefficients
data_table <- data.frame(lapply(df.coefficients, function(x) {
  if (is.numeric(x)) round(x, 3) else x
}))
T1 <- tableGrob(data_table, rows = NULL)

pdf(paste0(figure.path, "/lmm.coefficients_site.as.random.intercepts.pdf"), width = 6, height = 4)
print(p1)
dev.off()

pdf(paste0(figure.path, "/table.lmm.coefficients_site.as.random.intercepts.pdf"), width = 13, height = 13)
grid::grid.draw(T1)
dev.off()
```

**3.4 visualize model prediction**

I visualized the model predictions using sjPlot::plot_model(). uses the mean value of the moderator as well as one standard deviation below and above mean value to plot the effect of the moderator on the independent variable (following the convention suggested by Cohen and Cohen and popularized by Aiken and West (1991).
```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, fig.width=5, fig.height=5}
p.list1 <- list() # with confidence interval
p.list2 <- list() # without confidence interval
df.models.N.P.K2 <- df.models.N.P.K

# to make a range for the figures
df.range <- df.full %>%
  select(contains("var"), contains("change")) %>%
  gather(key = "sta.names") %>%
  na.omit()
df.range <- df.range %>% filter(!is.infinite(value))
df.range <- df.range %>%
  group_by(sta.names) %>%
  summarise(
    y.min = quantile(value, 0.05),
    y.max = quantile(value, 0.95)
  ) %>%
  ungroup()

df.models.N.P.K2 <- df.models.N.P.K2 %>% left_join(df.range)

# df.models.N.P.K2 %<>% filter(str_detect(sta.names, "Functional"))
for (i in 1:15) {
  mod <- df.models.N.P.K2$mod.out[[i]]
  p <- plot_model(mod, type = "pred", terms = c("log.pert [-2,2]", "end_year")) +
    theme_bw() +
    scale_color_manual(values = c("#56B4E9", "black", "#E69F00")) +
    scale_fill_manual(values = c("#56B4E9", "black", "#E69F00")) +
    # ylim(c(df.models.N.P.K2$y.min[[i]], df.models.N.P.K2$y.max[[i]])) +
    xlab(paste0("Perturbation (+", df.models.N.P.K2$trt[[i]], ")")) +
    ylab(df.models.N.P.K2$sta.names[[i]]) +
    ggtitle("") +
    theme(
      axis.text = element_text(color = "black"),
      panel.grid = element_blank()
    ) +
    ylim(-1.3, 1.3) +
    labs(color = "Year")
  p.list1[[i]] <- p
}
for (i in 1:15) {
  mod <- df.models.N.P.K2$mod.out[[i]]
  p <- plot_model(mod, type = "pred", terms = c("log.pert [-2,2]", "end_year"), se = FALSE) +
    theme_bw() +
    scale_color_manual(values = c("#56B4E9", "black", "#E69F00")) +
    scale_fill_manual(values = c("white", "white", "white")) +
    # ylim(c(df.models.N.P.K2$y.min[[i]], df.models.N.P.K2$y.max[[i]])) +
    xlab(paste0("Perturbation (+", df.models.N.P.K2$trt[[i]], ")")) +
    ylab(df.models.N.P.K2$sta.names[[i]]) +
    ggtitle("") +
    theme(
      axis.text = element_text(color = "black"),
      panel.grid = element_blank()
    ) +
    ylim(-1.3, 1.3) +
    labs(color = "Year")
  p.list2[[i]] <- p
}
```
First make a figure for the ones where the interaction between time and perturbation intensity is significant.

```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, fig.width=7, fig.height=3}
Fig <- wrap_plots(p.list1[[2]], p.list1[[4]], p.list1[[8]], nrow = 1, ncol = 3)
(Fig <- Fig +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom", legend.title = element_blank()))

pdf(paste0(figure.path, "/lmm.predictions.significant.interaction.pdf"), width = 7, height = 3)
Fig
dev.off()
```
Then make a figure for the rest predictions.
```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, fig.width=8, fig.height=6}
Fig <- wrap_plots(p.list2[[1]], p.list2[[3]], p.list2[[5]],
  p.list2[[6]], p.list2[[7]], p.list2[[9]], p.list2[[10]],
  p.list2[[11]], p.list2[[12]], p.list2[[13]], p.list2[[14]], p.list2[[15]],
  nrow = 3, ncol = 4
)
(Fig <- Fig +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom", legend.title = element_blank()))

pdf(paste0(getwd(), "/myfigure/lmm.predictions.non-significant.interaction.pdf"), width = 8, height = 6)
Fig
dev.off()
```
