---
title: "R codes for the NutNet-Stability project"
subtitle: "Step1: Prepare the data"
author: "Qiang Yang"
date: "`r Sys.Date()`"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{}
- \fancyfoot[CO,CE]{R codes for the NutNet-MultiStability project - prepare the data}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage



# Setup work environment

```{r, warning=FALSE, message=FALSE, highlight=T, prompt=TRUE, eval=T}
rm(list = ls())
gc()
# necessary R packages
Packages <- c(
  "tidyverse", "magrittr", "grid", "gridExtra", "rnaturalearth", "rnaturalearthdata",
  "ggcorrplot", "scales", "sp", "sf", "broom", "nlme", "scico", "vegan", "dlookr", "GGally",
  "PerformanceAnalytics"
)
pacman::p_load(char = Packages)
# functions with conflicting status (presence in different packages used here)
select <- dplyr::select
summarise <- dplyr::summarise
# working path
main.path <- getwd()
data.path <- paste0(main.path, "/mydata")
figure.path <- paste0(main.path, "/myfigure")
set.seed(1234)
```

\newpage

# Section 1: cover data cleaning 


**1.1 Get the raw data**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata <- read_csv(paste0(data.path, "/full-cover_2024-05-31.csv"))
```

**1.2 Exclude the records where live == zero** (i.e. plots that only include the dead part of the plants)

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata <- filter(mydata, live == 1)
```

**1.3 Check if there is any plot with treatment information missing**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
any(is.na(mydata$trt))
```

**1.4 Use the column site_code to represent site names**

There are some formatting issues with the column site_name, so I used the column site_code to represent site names.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata %<>% select(-site_name) # remove the site_name column
```

**1.5 Exclude the plots with any year from year0 to year4 data missing**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
plots.with.one.from.year0.to.year4.missing <- mydata %>%
  group_by(site_code, block, trt, plot) %>%
  nest() %>%
  mutate(result = map_lgl(data, function(dat) {
    length(setdiff(c(0, 1, 2, 3, 4), dat$year_trt)) > 0
  })) %>%
  filter(result) %>%
  select(-data, -result)

mydata <- anti_join(mydata, plots.with.one.from.year0.to.year4.missing)
rm(plots.with.one.from.year0.to.year4.missing)
```

**1.6 Now if there is a gap in sampling, delete the years after that gap, in case of multiple gaps, delete from the first gap**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata2 <- mydata %>%
  group_by(site_code, block, trt, plot) %>%
  nest()

mydata2 <- mydata2 %>%
  mutate(data2 = map(data, function(df1) {
    unique_years <- unique(df1$year_trt)
    if (length(unique_years) == max(unique_years) + 1) {
      df1 <- df1
    } else {
      gap_years <- setdiff(0:max(unique_years), unique_years)
      df1 <- df1 %>% filter(year_trt < min(gap_years))
    }
    return(df1)
  }))

mydata <- mydata2 %>%
  select(-data) %>%
  unnest(data2) %>%
  ungroup()
```


**1.7 Check how many subplots were sampled for each plot**

Here, I check if there are any plot, from which multiple subplots were sampled.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata %>%
  group_by(site_code, block, trt, plot) %>%
  summarise(n.subplot = length(unique(subplot))) %>%
  ungroup() %>%
  filter(n.subplot > 1)
```
Fine, no plot includes multiple subplots sampled.

**1.8 Exclude the blocks that only with one treatment (including control)**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
single.treatment.block <- mydata %>%
  group_by(site_code, block) %>%
  summarize(n.trt = n_distinct(trt)) %>%
  ungroup() %>%
  filter(n.trt == 1)
mydata <- anti_join(mydata, single.treatment.block) # removed the relevant plots
rm(single.treatment.block)
```

**1.9 Exclude the blocks that without control plot**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
no.control.block <- mydata %>%
  group_by(site_code, block) %>%
  summarize(control.check = any(unique(trt) == "Control")) %>%
  ungroup() %>%
  filter(!control.check)
mydata %<>% anti_join(no.control.block)
```

**1.10 Compile the cleaned cover data**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata.coverage <- mydata %>%
  droplevels() %>%
  distinct()
rm(mydata)
```

# Section 2: Harmonize taxonomic names in the cover dataset

In this section, taxonomic names **within each site** were harmonized.

**2.1 first-round standardization using the codes written by Yann**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
data1 <- mydata.coverage
code.path <- paste0(main.path, "/mycode")
source(paste0(code.path, "/Taxonomic.Adjustments.function.191115.R"))
library(plyr)
data1 <- Taxonomic.Adjustments(datafile = data1)
```

**2.2 remove the taxa UNKNOWN**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
data1$Taxon %>%
  str_subset("UNKNOWN") %>%
  unique()
data1 %<>% filter(!str_detect(Taxon, "UNKNOWN"))
```

**2.3 aggregate the taxa of a genus to genus sp.** when an unknown species of the genus exist.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
aggregate.to.genus.level.for.sp.f <- function(Site.code = "bldr.us") {
  df <- data1 %>%
    filter(site_code == Site.code)
  df <- df %>%
    mutate(splitted = map(
      Taxon,
      function(x) str_split(x, pattern = " ")[[1]]
    )) %>%
    mutate(
      genus = map_chr(splitted, function(x) x[1]),
      epithet = map_chr(splitted, function(x) x[2])
    )
  genus.to.aggregate <- df %>%
    filter(epithet == "SP.") %>%
    select(genus) %>%
    distinct() %>%
    unlist()
  if (length(genus.to.aggregate) > 0) {
    df1 <- df %>%
      filter(genus %in% genus.to.aggregate)
    df2 <- df %>%
      filter(genus %in% setdiff(df$genus, genus.to.aggregate))
    df1 %<>% mutate(Taxon = paste(genus, "SP."))
    df <- df1 %>% bind_rows(df2)
    df %<>% select(-splitted, -genus, -epithet)
    df %<>%
      dplyr::group_by(
        site_code, block, plot, subplot,
        trt, year, year_trt, Taxon
      ) %>%
      dplyr::summarise(max_cover = sum(max_cover)) %>%
      ungroup()
  } else {
    df %<>%
      select(-splitted, -genus, -epithet) %>%
      as_tibble()
  }
  return(df)
}

L <- data1 %>%
  select(site_code) %>%
  distinct() %>%
  as_tibble() %>%
  mutate(dat = map(site_code, aggregate.to.genus.level.for.sp.f))
data1 <- L %>%
  dplyr::rename(sites = site_code) %>%
  unnest(cols = dat) %>%
  select(-sites)

mydata.coverage <- data1
rm(data1)
detach("package:plyr", unload = TRUE)
save(mydata.coverage, file = paste0(data.path, "/mydata.coverage.Rdata"))
```

# Section 3: biomass data cleaning 

**3.1 Get the raw data**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata <- read_csv(paste0(data.path, "/full-biomass_2024-05-31.csv"))
```

**3.2 Exclude the records where live == zero** (i.e. plots that only include the dead part of the plants)

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata <- filter(mydata, live == 1)
```

**3.3 Check if there is any plot with treatment information missing**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
any(is.na(mydata$trt))
```

**3.4 Use the column site_code to represent site names**

There are some formatting issues with the column site_name, so I used the column site_code to represent site names.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata %<>% select(-site_name) # remove the site_name column
```

**3.5 Exclude the plots with any year from year0 to year4 data missing**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
plots.with.one.from.year0.to.year4.missing <- mydata %>%
  group_by(site_code, block, trt, plot) %>%
  nest() %>%
  mutate(result = map_lgl(data, function(dat) {
    length(setdiff(c(0, 1, 2, 3, 4), dat$year_trt)) > 0
  })) %>%
  filter(result) %>%
  select(-data, -result)

mydata <- anti_join(mydata, plots.with.one.from.year0.to.year4.missing)
rm(plots.with.one.from.year0.to.year4.missing)
```

**3.6 Now if there is a gap in sampling, delete the years after that gap, in case of multiple gaps, delete from the first gap**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata2 <- mydata %>%
  group_by(site_code, block, trt, plot) %>%
  nest()

mydata2 <- mydata2 %>%
  mutate(data2 = map(data, function(df1) {
    unique_years <- unique(df1$year_trt)
    if (length(unique_years) == max(unique_years) + 1) {
      df1 <- df1
    } else {
      gap_years <- setdiff(0:max(unique_years), unique_years)
      df1 <- df1 %>% filter(year_trt < min(gap_years))
    }
    return(df1)
  }))

mydata <- mydata2 %>%
  select(-data) %>%
  unnest(data2) %>%
  ungroup()
```


**3.7 Check how many subplots were sampled for each plot**

Here, I check if there are any plot, from which multiple subplots were sampled.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata %>%
  group_by(site_code, block, trt, plot) %>%
  summarise(n.subplot = length(unique(subplot))) %>%
  ungroup() %>%
  filter(n.subplot > 1)
```
Fine, no plot includes multiple subplots sampled.

**3.8 Exclude the blocks that only with one treatment (including control)**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
single.treatment.block <- mydata %>%
  group_by(site_code, block) %>%
  summarize(n.trt = n_distinct(trt)) %>%
  ungroup() %>%
  filter(n.trt == 1)
mydata <- anti_join(mydata, single.treatment.block) # removed the relevant plots
rm(single.treatment.block)
```

**3.9 Exclude the blocks that without control plot**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
no.control.block <- mydata %>%
  group_by(site_code, block) %>%
  summarize(control.check = any(unique(trt) == "Control")) %>%
  ungroup() %>%
  filter(!control.check)
mydata %<>% anti_join(no.control.block)
```

**3.10 Compile the cleaned biomass data**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata.biomass <- mydata %>%
  droplevels() %>%
  distinct()
save(mydata.biomass, file = paste0(data.path, "/mydata.biomass.Rdata"))
```

# Section 4: merge coverage data with biomass data

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
mydata.coverage.nested <- mydata.coverage %>%
  group_by(site_code, block, trt, plot) %>%
  nest() %>%
  rename(coverage.data = data)
mydata.biomass.nested <- mydata.biomass %>%
  group_by(site_code, block, trt, plot) %>%
  nest() %>%
  rename(biomass.data = data)
df.tidy <- mydata.coverage.nested %>%
  full_join(mydata.biomass.nested)
# now add indicator for the availability of coverage or biomass for each plot
df.tidy <- df.tidy %>%
  mutate(data.availability = map2_chr(coverage.data, biomass.data, function(df1, df2) {
    if (is.null(df1)) {
      output <- "only biomass data"
    } else
    if (is.null(df2)) {
      output <- "only coverage data"
    } else {
      output <- "both biomass and coverage data"
    }
    return(output)
  }))
df.tidy <- df.tidy %>%
  ungroup()
```


# Section 5: add plot characteristics

In this section, I compile plot characteristics including total coverage, total biomass, species richness, evenness at year0, geographic coordinates (site level), and original nutrient concentration.

**5.1 total coverage, total biomass, species richness, evenness at year0**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}

df.tidy <- df.tidy %>%
  mutate(total.cover = map2_dbl(coverage.data, data.availability, function(dat, DA) {
    if (DA == "only biomass data") {
      out <- NA
    } else {
      out <- dat %>%
        filter(year_trt == 0) %>%
        pull(max_cover) %>%
        sum()
    }
    return(out)
  })) %>%
  mutate(richness = map2_dbl(coverage.data, data.availability, function(dat, DA) {
    if (DA == "only biomass data") {
      out <- NA
    } else {
      out <- dat %>%
        filter(year_trt == 0) %>%
        pull(Taxon) %>%
        n_distinct()
    }
    return(out)
  })) %>%
  mutate(shannon.diversity = map2_dbl(coverage.data, data.availability, function(dat, DA) {
    if (DA == "only biomass data") {
      out <- NA
    } else {
      out <- dat %>%
        filter(year_trt == 0) %>%
        pull(max_cover) %>%
        vegan::diversity()
    }
    return(out)
  })) %>%
  mutate(total.mass = map2_dbl(biomass.data, data.availability, function(dat, DA) {
    if (DA == "only coverage data") {
      out <- NA
    } else {
      out <- dat %>%
        filter(year_trt == 0) %>%
        pull(mass) %>%
        sum()
    }
    return(out)
  })) %>%
  mutate(pielou.evenness = shannon.diversity / log(richness))
# set the pielou evenness of the plots that only have one species to 0
df.tidy$pielou.evenness[df.tidy$richness == 1] <- 0
```

**5.2 add initial nutrient concentration for each plot**
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.nutrient <-
  read_csv(paste0(data.path, "/full-soil_2024-05-31.csv")) %>%
  rename(trt = Treatment) %>%
  select(site_code, block, plot, year_trt, trt, pct_N, ppm_P, ppm_K) %>%
  filter(year_trt == 0)
# replace the ‘NULL’ in the rest columns with ‘NA’
replace.null.f <- function(x) {
  x[x == "NULL"] <- NA
  y <- as.numeric(x)
  return(y)
}
df.nutrient %<>% mutate_at(c("pct_N", "ppm_P", "ppm_K"), replace.null.f)
# remove the blocks if all the plots in the block have no data for one nutrient
blocks.to.exclude <- df.nutrient %>%
  group_by(site_code, block) %>%
  nest() %>%
  mutate(check_ = map_lgl(data, function(dat) {
    all(is.na(dat$pct_N)) | all(is.na(dat$ppm_P)) | all(is.na(dat$ppm_K))
  })) %>%
  filter(check_) %>%
  ungroup() %>%
  select(site_code, block)
df.nutrient %<>% anti_join(blocks.to.exclude)

# I noticed that some plots have no nutrient information,
# but most of the other plots in the block has nutrient information.
# The mean value of these plots were assigned to the plots with missing value
df.nutrient %<>%
  group_by(site_code, block) %>%
  nest() %>%
  mutate(data = map(data, imputeTS::na_mean)) %>%
  unnest(cols = data) %>%
  ungroup()
# add three columns in the dataset to indicate the addition of N, P, K
df.nutrient.addition <- data.frame(
  trt = c("Control", "N", "P", "K", "NP", "NK", "PK", "NPK", "Fence", "NPK+Fence"),
  N_addition = c(0, 1, 0, 0, 1, 1, 0, 1, 0, 1),
  P_addition = c(0, 0, 1, 0, 1, 0, 1, 1, 0, 1),
  K_addition = c(0, 0, 0, 1, 0, 1, 1, 1, 0, 1)
)
df.nutrient <- left_join(df.nutrient, df.nutrient.addition)
# add three columns in the dataset to indicate the ratio of the added nutrients
df.nutrient %<>%
  mutate(
    N_added_proportion = N_addition * 10 / pct_N,
    P_added_proportion = P_addition * 10 / ppm_P,
    K_added_proportion = K_addition * 10 / ppm_K
  )
# add the nutrient data
df.tidy %<>% inner_join(df.nutrient %>% select(-year_trt))
# have a look at the missing value
df.tidy %>%
  select(-coverage.data, -biomass.data) %>%
  diagnose()
# good. NA only exist for the plot with only coverage or biomass data
```

**5.3 further cleaning**

It is possible that some blocks lose all control plots during the section above. Remove such blocks.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
no.control.block <- df.tidy %>%
  group_by(site_code, block) %>%
  summarize(control.check = any(unique(trt) == "Control")) %>%
  filter(!control.check)
df.tidy %<>%
  anti_join(no.control.block, by = c("site_code", "block"))
```

For this project, only keep the plots with the Control, +N, +P, and +K treatments.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.tidy <- df.tidy %>% 
  filter(trt %in% c("Control", "N", "P", "K"))
```

**5.4 geographic coordinates for each site and make a map**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
all.site.location <- read_csv(paste0(data.path, "/sites_2023-11-07.csv"))
setdiff(df.tidy$site_code, all.site.location$site_code)
# all sites have location information
df.tidy <- df.tidy %>%
  left_join(all.site.location %>%
    select(site_code, latitude, longitude))
rm(all.site.location)
save(df.tidy, file = paste0(data.path, "/df.tidy.Rdata"))
```

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
load(file = paste0(data.path, "/df.tidy.Rdata"))
sites <- df.tidy %>% 
  select(site_code, latitude, longitude) %>% 
  distinct()

# Load world map
world <- ne_countries(scale = "medium", returnclass = "sf")

# Convert site data to an sf object
sites_sf <- st_as_sf(sites, coords = c("longitude", "latitude"), crs = 4326)

# Create the map
(
  p.map <- ggplot(data = world) +
    geom_sf(fill = "gray90", color = "gray40", size = 0.1) + # Clean world map
    geom_sf(data = sites_sf, color = "white", size = 1.5, shape = 21, fill = "red", stroke = 0.2) + # Site points
    theme_minimal() +
    theme(
      panel.background = element_rect(fill = "white", color = NA), # Clean background
      panel.grid.major = element_blank(), # Remove major grid lines
      panel.grid.minor = element_blank(), # Remove minor grid lines
      axis.text = element_blank(), # Remove axis text
      axis.title = element_blank(), # Remove axis titles
      axis.ticks = element_blank(), # Remove axis ticks
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      plot.caption = element_text(hjust = 0.5, size = 10)
    )
)
ggsave(
  filename = paste0(figure.path, paste("/Locations of the", n_distinct(sites), "selected sites.pdf")),
  width = 4,
  height = 3
)
```
