---
title: "R codes for the NutNet-Stability project"
subtitle: "Step4.2: Dimensionality analysis following discussion on 09.01.2025"
author: "Qiang Yang"
date: "`r Sys.Date()`"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{}
- \fancyfoot[CO,CE]{R codes for the NutNet-stability project}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

I will test the significance of the difference in the length of the first semi-axis of the confidence ellipsoid of the covariance matrix between the high-perturbation group and the low-perturbation group. The analysis involves the following steps:

- Calculate Semi-Axis Lengths: Compute the semi-axis length of the confidence ellipsoid for each time point (e.g., year) in both the high-perturbation and low-perturbation groups across each of the three nutrient treatments. This will result in six time series (two groups × three treatments), as illustrated in the figure "decoupling.var.explained_func_var_det.pdf".

- Compare Mean Semi-Axis Lengths: For each treatment, compare the mean semi-axis lengths of the two time series (high- and low-perturbation groups). Since the time series are likely auto-correlated, a simple pairwise t-test is not appropriate. Instead, use a generalized least squares (GLS) model. In the GLS model, treat perturbation intensity (high vs. low) as the grouping factor and extract the model coefficient representing the difference between the two perturbation levels.

- Construct Null Distributions: To test the significance of the observed difference, construct null distributions of the GLS model coefficient:
     a. randomly assign plots into high- and low-perturbation groups.
     b. Repeat the GLS process described in step 2 for these randomized groupings.
     c. Perform this randomization multiple times to build a null distribution of the GLS model coefficient under the assumption of no true difference.

- Test Significance: Compare the observed GLS model coefficient (from the actual data) against the null distribution of GLS model coefficients. The observed value's position within the null distribution will determine the statistical significance of the difference.

- Repeat for Each Treatment: Conduct the above analysis separately for each of the three nutrient treatments to assess treatment-specific differences.

- Note that the codes does not exactly follow the steps mentioned above as I will try to make the codes in the tidy manner so some of the process will be performed in a vectorised way.


# Section 1: setup work environment

```{r, warning=FALSE, message=FALSE, highlight=T, eval=T}
rm(list = ls())
gc()
# necessary R packages
Packages <- c(
  "tidyverse", "magrittr", "grid", "gridExtra", "scales",
  "broom", "dlookr", "nlme", "patchwork", "cowplot"
)
pacman::p_load(char = Packages)
# functions with conflicting status (presence in different packages used here)
select <- dplyr::select
summarise <- dplyr::summarise
# working path
main.path <- getwd()
data.path <- paste0(main.path, "/mydata")
figure.path <- paste0(main.path, "/myfigure")
set.seed(1234)
```



# Section 2: dimensionality analysis using the PCA approach (func_var_det used)

There are two functional variability. Here I will use func_var_det. But I will use func_var in section 3 to check if conclusion will be changed.

**2.1 data organization - observed data**

Note that all stability components here are first scaled so that they have the similar scale.
```{r, warning=FALSE, message=FALSE, eval=F, highlight=T, fig.width=7, fig.height=4}
load(paste0(data.path, "/df.tidy.stabilities.Rdata"))
load(paste0(data.path, "/df.tidy.Rdata"))
df.full <- df.tidy.stabilities %>%
  left_join(df.tidy %>% select(-contains("data")))
rm(df.tidy, df.tidy.stabilities)

df.full %<>%
  select(site_code:end_year, matches("_added_proportion|change|var"))
sta.names <- colnames(df.full) %>% str_subset("change|var")

# For the PCA analysis, all stability components need to be available for each row
# no NA, no infinite
# remove the relevant cases
df.full <- df.full %>%
  filter_all(all_vars(!is.infinite(.))) %>%
  filter_all(all_vars(!is.na(.)))

df.stability <- df.full

# scale stability components
df.stability %<>% mutate(across(one_of(sta.names), scale))


# make a meta data for multidimensionality analysis
trt.set <- c("N", "P", "K")
final.year.set <- 9 # original using 5:9
df.meta <- crossing(trt.set, final.year.set)

## add data, nested by year

df.meta <- df.meta %>%
  mutate(data = map2(
    trt.set, final.year.set,
    function(x1, x2) {
      plots.to.keep <- df.stability %>%
        filter(
          trt == x1,
          end_year == x2
        ) %>%
        select(site_code, block, trt, plot)
      subdat <- df.stability %>%
        semi_join(plots.to.keep) %>%
        filter(
          trt == x1,
          end_year <= x2
        )
      return(subdat)
    }
  ))
df.meta %<>% ungroup()

## add data, nested by perturbation intensity and end_year

df.meta <- df.meta %>%
  mutate(data2 = map2(
    trt.set, data,
    function(x1, dat) {
      df.pert <- dat %>%
        select(site_code, block, trt, plot, paste0(x1, "_added_proportion")) %>%
        distinct() %>%
        set_colnames(c("site_code", "block", "trt", "plot", "PI"))
      df.pert %<>%
        mutate(PI.group = ifelse(PI > median(PI), "High", "Low"))
      dat %<>% left_join(df.pert)
      dat.nested <- dat %>%
        group_by(end_year, PI.group) %>%
        nest() %>%
        ungroup()
      return(dat.nested)
    }
  ))

df.meta %<>%
  select(-data) %>%
  unnest(cols = data2)

df.N.P.K <- df.meta %>% ungroup()

# # further only select stability components from data
df.N.P.K %<>%
  mutate(data = map(data, function(dat) dat %>% select(one_of(sta.names))))
df.N.P.K.obs <- df.N.P.K
```

**2.2 data organization - data produced by permutation**

Conduct permutations. The dataset full were permuted by plot*final.year.set*end_year.

```{r, warning=FALSE, message=FALSE, eval=F, highlight=T, fig.width=7, fig.height=4}
set.seed(1234)
L <- list()
for (i in 1:1000) {
  # Specify the column names to permute
  print(i)
  print(Sys.time())
  df.stability.obs <- df.stability
# first randomize the nutrient perturbaiton intensity of plots
df.nutrient.pert <- df.stability.obs %>% 
  select(site_code, block, trt, plot, contains("proportion")) %>% 
  distinct()# the nutrient perturbation of each plot
df.nutrient.pert.random <- df.nutrient.pert %>% 
  group_by(trt) %>% 
  nest() %>% 
  mutate(data2 = map(data, function(dt_){
    dt_ %>% 
      mutate(N_added_proportion = sample(N_added_proportion),
             P_added_proportion = sample(P_added_proportion),
             K_added_proportion = sample(K_added_proportion))
  })) %>% 
  select(-data) %>% 
  unnest(cols = data2)
df.stability.random.plots <- df.stability.obs %>% 
  select(-contains("proportion")) %>% 
  left_join(df.nutrient.pert.random)

df.stability.0 <- df.stability.random.plots
rm(df.stability.random.plots)

  # make a meta data for multidimensionality analysis
  trt.set <- c("N", "P", "K")
  final.year.set <- 9 # original using 5:9
  df.meta <- crossing(trt.set, final.year.set)

  ## add data, nested by year

  df.meta <- df.meta %>%
    mutate(data = map2(
      trt.set, final.year.set,
      function(x1, x2) {
        plots.to.keep <- df.stability.0 %>%
          filter(
            trt == x1,
            end_year == x2
          ) %>%
          select(site_code, block, trt, plot)
        subdat <- df.stability.0 %>%
          semi_join(plots.to.keep) %>%
          filter(
            trt == x1,
            end_year <= x2
          )
        return(subdat)
      }
    ))
  df.meta %<>% ungroup()

  ## add data, nested by perturbation intensity and end_year

  df.meta <- df.meta %>%
    mutate(data2 = map2(
      trt.set, data,
      function(x1, dat) {
        df.pert <- dat %>%
          select(site_code, block, trt, plot, paste0(x1, "_added_proportion")) %>%
          distinct() %>%
          set_colnames(c("site_code", "block", "trt", "plot", "PI"))
        df.pert %<>%
          mutate(PI.group = ifelse(PI > median(PI), "High", "Low"))
        dat %<>% left_join(df.pert)
        dat.nested <- dat %>%
          group_by(end_year, PI.group) %>%
          nest() %>%
          ungroup()
        return(dat.nested)
      }
    ))

  df.meta %<>%
    select(-data) %>%
    unnest(cols = data2)

  df.N.P.K <- df.meta %>% ungroup()

  # # further only select stability components from data
  df.N.P.K %<>%
    mutate(data = map(data, function(dat) dat %>% select(one_of(sta.names))))

  L[[i]] <- df.N.P.K
  rm(df.N.P.K)
}

names(L) <- 1:1000
df.N.P.K.perm <- tibble(
  perm.id = names(L), # This will be the ID column
  data = L # This will be the column containing data.frames
) %>%
  unnest(data) %>%
  mutate(perm.id = as.numeric(perm.id))

df.permutation <- df.N.P.K.obs %>%
  mutate(
    data.group = "obs",
    perm.id = 0
  ) %>%
  bind_rows(df.N.P.K.perm %>%
    mutate(data.group = "permutation"))
save(df.permutation, file = paste0(data.path, "/df.permutation.random.plots.Rdata"))
```

**2.3 compute the covariance matrix and the eigen vector of the covariance matrix** 

Here I will choose func_var_det.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
load(file = paste0(data.path, "/df.permutation.random.plots.Rdata"))
df.permutation <- df.permutation %>%
  mutate(data = map(data, function(dt1) dt1 %>% select(-func_var)))
# calculate covariance matrix and eigenvector
df.permutation %<>% mutate(cov.M = map(data, cov))
df.permutation %<>% mutate(eigvec = map(cov.M, function(x) eigen(x)$vectors))
#### calculate the variance explained by first eigen vector of the covariance matrix
df.permutation %<>%
  mutate(var.explained = map2_dbl(data, eigvec, function(D, X) {
    D <- as.matrix(D)
    P <- D %*% X
    V <- apply(P, 2, var)
    V[1] / sum(V)
  }))

df.variance.explained <- df.permutation %>%
  select(trt.set, final.year.set, end_year, PI.group, data.group, perm.id, var.explained)
```

**2.4 a figure showing the obs of the change in dimensionality** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=5, fig.height=5}
p.var.explained <-
  df.variance.explained %>%
  filter(final.year.set == 9, perm.id == 0) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("N", "P", "K"))) %>%
  ggplot(aes(x = end_year, y = var.explained, color = trt.set, linetype = PI.group)) +
  geom_line(linewidth = 1) +
  # geom_point(aes(shape = significance), size = 1.5) +
  scale_color_brewer(
    name = "Treatment:",
    palette = "Dark2",
    labels = c("+N", "+P", "+K")
  ) +
  # scale_shape_manual(values = c(4, 19)) +
  theme_bw() +
  theme(
    legend.position = "top", legend.box = "vertical", legend.margin = margin(),
    axis.text = element_text(color = "black", size = 10),
    panel.grid = element_blank()
  ) +
  xlab("Years") +
  ylab("Relative length of the first semi-axis") +
  labs(linetype = "Perturbation:") +
  guides(
    color = guide_legend(nrow = 1), # Arrange color legend in 1 row
    linetpye = guide_legend(nrow = 1) # Arrange shape legend in 1 row
  )
pdf(paste0(getwd(), "/myfigure/decoupling.var.explained_func_var_det.pdf"), width = 5.5, height = 5.5)
p.var.explained
dev.off()
p.var.explained
p.decoupling.var.explained_func_var_det <- p.var.explained
```

**2.5 GLS for the observed and the permutated** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=5, fig.height=4}
# i first nest the df.variance explained, so that each row can have one gls
df.nested <- df.variance.explained %>%
  select(-final.year.set) %>% # unique value, so can be removed
  group_by(trt.set, perm.id, data.group) %>%
  nest()
# perform the gls
df.nested <- df.nested %>%
  mutate(model.results = map(data, function(dt_) {
    gls_model <- gls(var.explained ~ PI.group,
      data = dt_,
      correlation = corAR1(form = ~ end_year | PI.group)
    )
  }))
# extract the coefficient representing the difference in semi-axis length
df.nested <- df.nested %>%
  mutate(diff_value = map_dbl(model.results, function(model_) coef(model_)["PI.groupLow"]))

# note that the difference now is calculated as low-perturbation minus high-perturbation,
# I want to use it in an opposite way
df.nested <- df.nested %>%
  mutate(diff_value = -1 * diff_value)
# now factorize trt.set
df.nested <- df.nested %>%
  mutate(trt.set = paste0("+", trt.set)) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("+N", "+P", "+K")))

# only keep the key columns
df.results <- df.nested %>%
  select(-data, -model.results) %>%
  group_by(trt.set) %>%
  nest()
# now add a column to represent the significance
df.results <- df.results %>%
  mutate(alpha = map_dbl(data, function(dt_) {
    x <- dt_ %>%
      filter(perm.id != 0) %>%
      select(diff_value) %>%
      unlist()
    x0 <- dt_ %>%
      filter(perm.id == 0) %>%
      select(diff_value) %>%
      unlist()
    length(which(x <= x0)) / length(x)
  }))
# also add a denote colomn for p values so I can add it to the figure later
df.results <- df.results %>%
  mutate(denote = map_chr(alpha, function(x) {
    paste("P =", x)
  }))
# add a column represent significance groups
df.results <- df.results %>%
  mutate(sig.group = map_chr(alpha, function(x) {
    if (x < 0.05) {
      y <- "negative"
    } else if (x < 0.95) {
      y <- "non-significant"
    } else {
      y <- "positive"
    }
  }))

# now make a figure

p.null <- df.nested %>%
  filter(data.group == "permutation") %>%
  ggplot(aes(x = diff_value)) +
  facet_wrap(~trt.set, nrow = 3) +
  geom_histogram(bins = 50, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% filter(data.group == "obs") %>% left_join(df.results %>% select(trt.set, sig.group)),
    aes(xintercept = diff_value, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% mutate(x = -0.13, y = 55),
    aes(x = x, y = y, label = denote)
  ) +
  theme_bw() +
  xlab("Difference in the relative length\n of the first semi-axis") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "top",
    panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black")) +
  labs(color = "Significance:")


pdf(paste0(getwd(), "/myfigure/decoupling.figure.stats_null.model_func_var_det.pdf"), width = 5, height = 5)
p.null
dev.off()
p.null
df.nested.decoupling.var.explained_func_var_det <- df.nested
df.results.decoupling.var.explained_func_var_det <- df.results
p.null.decoupling.var.explained_func_var_det <- p.null
```

**2.6 combine the pattern figure with the null model figure** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=8, fig.height=6}
p.combined <- (p.decoupling.var.explained_func_var_det | plot_spacer() | p.null.decoupling.var.explained_func_var_det) +
  plot_layout(widths = c(1, 0.05, 1)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(face = "bold"))
pdf(paste0(getwd(), "/myfigure/SI.decoupling.figure.func_var_det.pdf"), width = 8, height = 6)
p.combined
dev.off()
```

**2.7 make a figure only showing the main effect of perturbation** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=3, fig.height=6}
df.pert.effect <- df.nested.decoupling.var.explained_func_var_det %>%
  filter(data.group == "obs") %>%
  mutate(
    high = map_dbl(model.results, function(model_) coef(model_)[1]),
    low = map_dbl(model.results, function(model_) coef(model_)[1] + coef(model_)[2])
  ) %>%
  ungroup()
df.pert.effect.long <- df.pert.effect %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, high, low) %>%
  pivot_longer(cols = c(high, low), names_to = "Perturbation", values_to = "Value")

df.pvalues <- df.results.decoupling.var.explained_func_var_det %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, denote) %>%
  arrange(desc(Treatment))
df.pvalues$x <- rep(1, 3)
df.pvalues$y <- c(0.38, 0.4, 0.42)

df.pert.effect.long <- df.pert.effect.long %>% 
  left_join(df.results.decoupling.var.explained_func_var_det %>%
              rename(Treatment = trt.set) %>% select(Treatment, alpha))
df.pert.effect.long <- df.pert.effect.long %>% 
  mutate(Significance = map_chr(alpha, function(x)ifelse(x<0.05|x>0.95, "significant","non-significant")))

# later decided not to add pvalues but use linetype to represent significance

P.decoupling.main_func_var_det <- df.pert.effect.long %>%
  mutate(Perturbation = as.factor(Perturbation)) %>%
  mutate(Perturbation = fct_relevel(Perturbation, c("low", "high"))) %>%
  ggplot(aes(x = as.numeric(Perturbation), y = Value, color = Treatment)) +
  geom_point(size = 2.5) +
  geom_line(linewidth = 1, aes(linetype = Significance)) +
  #geom_text(data = df.pvalues, aes(x = x, y = y, label = denote, color = Treatment), hjust = 0, show.legend = F) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  scale_x_continuous(breaks = c(1, 2), labels = c("low", "high")) +
  xlab("Perturbation") +
  ylab("Relative length of the first semi-axis") +
  theme(axis.text = element_text(color = "black"),
        panel.grid = element_blank())+
  scale_linetype_manual(values = c("dashed","solid"))
```


# Section 3: dimensionality analysis using the PCA approach (func_var used)


**3.1 compute the covariance matrix and the eigen vector of the covariance matrix** 

Here I will choose func_var.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
load(file = paste0(data.path, "/df.permutation.random.plots.Rdata"))
df.permutation <- df.permutation %>%
  mutate(data = map(data, function(dt1) dt1 %>% select(-func_var_det)))
# calculate covariance matrix and eigenvector
df.permutation %<>% mutate(cov.M = map(data, cov))
df.permutation %<>% mutate(eigvec = map(cov.M, function(x) eigen(x)$vectors))
#### calculate the variance explained by first eigen vector of the covariance matrix
df.permutation %<>%
  mutate(var.explained = map2_dbl(data, eigvec, function(D, X) {
    D <- as.matrix(D)
    P <- D %*% X
    V <- apply(P, 2, var)
    V[1] / sum(V)
  }))

df.variance.explained <- df.permutation %>%
  select(trt.set, final.year.set, end_year, PI.group, data.group, perm.id, var.explained)
```

**3.2 a figure showing the obs of the change in dimensionality** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=6, fig.height=6}
p.var.explained <-
  df.variance.explained %>%
  filter(final.year.set == 9, perm.id == 0) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("N", "P", "K"))) %>%
  ggplot(aes(x = end_year, y = var.explained, color = trt.set, linetype = PI.group)) +
  geom_line(linewidth = 1) +
  # geom_point(aes(shape = significance), size = 1.5) +
  scale_color_brewer(
    name = "Treatment:",
    palette = "Dark2",
    labels = c("+N", "+P", "+K")
  ) +
  # scale_shape_manual(values = c(4, 19)) +
  theme_bw() +
  theme(
    legend.position = "top", legend.box = "vertical", legend.margin = margin(),
    axis.text = element_text(color = "black", size = 10),
    panel.grid = element_blank()
  ) +
  xlab("Years") +
  ylab("Relative length of the first semi-axis") +
  labs(linetype = "Perturbation:") +
  guides(
    color = guide_legend(nrow = 1), # Arrange color legend in 1 row
    linetpye = guide_legend(nrow = 1) # Arrange shape legend in 1 row
  )
pdf(paste0(getwd(), "/myfigure/decoupling.var.explained_func_var.pdf"), width = 5.5, height = 5.5)
p.var.explained
dev.off()
p.var.explained
p.decoupling.var.explained_func_var <- p.var.explained
```

**3.3 a figure showing the stats of the change in dimensionality** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=5, fig.height=5}
# i first nest the df.variance explained, so that each row can have one gls
df.nested <- df.variance.explained %>%
  select(-final.year.set) %>% # unique value, so can be removed
  group_by(trt.set, perm.id, data.group) %>%
  nest()
# perform the gls
df.nested <- df.nested %>%
  mutate(model.results = map(data, function(dt_) {
    gls_model <- gls(var.explained ~ PI.group,
      data = dt_,
      correlation = corAR1(form = ~ end_year | PI.group)
    )
  }))
# extract the coefficient representing the difference in semi-axis length
df.nested <- df.nested %>%
  mutate(diff_value = map_dbl(model.results, function(model_) coef(model_)["PI.groupLow"]))

# note that the difference now is calculated as low-perturbation minus high-perturbation,
# I want to use it in an opposite way
df.nested <- df.nested %>%
  mutate(diff_value = -1 * diff_value)
# now factorize trt.set
df.nested <- df.nested %>%
  mutate(trt.set = paste0("+", trt.set)) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("+N", "+P", "+K")))

# only keep the key columns
df.results <- df.nested %>%
  select(-data, -model.results) %>%
  group_by(trt.set) %>%
  nest()
# now add a column to represent the significance
df.results <- df.results %>%
  mutate(alpha = map_dbl(data, function(dt_) {
    x <- dt_ %>%
      filter(perm.id != 0) %>%
      select(diff_value) %>%
      unlist()
    x0 <- dt_ %>%
      filter(perm.id == 0) %>%
      select(diff_value) %>%
      unlist()
    length(which(x <= x0)) / length(x)
  }))
# also add a denote colomn for p values so I can add it to the figure later
df.results <- df.results %>%
  mutate(denote = map_chr(alpha, function(x) {
    paste("P =", x)
  }))
# add a column represent significance groups
df.results <- df.results %>%
  mutate(sig.group = map_chr(alpha, function(x) {
    if (x < 0.05) {
      y <- "negative"
    } else if (x < 0.95) {
      y <- "non-significant"
    } else {
      y <- "positive"
    }
  }))

# now make a figure

p.null <- df.nested %>%
  filter(data.group == "permutation") %>%
  ggplot(aes(x = diff_value)) +
  facet_wrap(~trt.set, nrow = 3) +
  geom_histogram(bins = 50, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% filter(data.group == "obs") %>% left_join(df.results %>% select(trt.set, sig.group)),
    aes(xintercept = diff_value, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% mutate(x = -0.13, y = 55),
    aes(x = x, y = y, label = denote)
  ) +
  theme_bw() +
  xlab("Difference in the relative length\n of the first semi-axis") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "top",
    panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black")) +
  labs(color = "Significance:")


pdf(paste0(getwd(), "/myfigure/decoupling.figure.stats_null.model_func_var.pdf"), width = 5, height = 5)
p.null
dev.off()
p.null
df.nested.decoupling.var.explained_func_var <- df.nested
df.results.decoupling.var.explained_func_var <- df.results
p.null.decoupling.var.explained_func_var <- p.null
```

**3.4 combine the pattern figure with the null model figure** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=8, fig.height=6}
p.combined <- (p.decoupling.var.explained_func_var | plot_spacer() | p.null.decoupling.var.explained_func_var) +
  plot_layout(widths = c(1, 0.05, 1)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(face = "bold"))
pdf(paste0(getwd(), "/myfigure/SI.decoupling.figure.func_var.pdf"), width = 8, height = 6)
p.combined
dev.off()
```
Seems conclusions not changed.

**3.5 make a figure only showing the main effect of perturbation** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=3, fig.height=6}
df.pert.effect <- df.nested.decoupling.var.explained_func_var %>%
  filter(data.group == "obs") %>%
  mutate(
    high = map_dbl(model.results, function(model_) coef(model_)[1]),
    low = map_dbl(model.results, function(model_) coef(model_)[1] + coef(model_)[2])
  ) %>%
  ungroup()
df.pert.effect.long <- df.pert.effect %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, high, low) %>%
  pivot_longer(cols = c(high, low), names_to = "Perturbation", values_to = "Value")

df.pvalues <- df.results.decoupling.var.explained_func_var %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, denote) %>%
  arrange(desc(Treatment))
df.pvalues$x <- rep(1, 3)
df.pvalues$y <- c(0.38, 0.4, 0.42)

df.pert.effect.long <- df.pert.effect.long %>% 
  left_join(df.results.decoupling.var.explained_func_var %>%
              rename(Treatment = trt.set) %>% select(Treatment, alpha))
df.pert.effect.long <- df.pert.effect.long %>% 
  mutate(Significance = map_chr(alpha, function(x)ifelse(x<0.05|x>0.95, "significant","non-significant")))

# later decided not to add pvalues but use linetype to represent significance

P.decoupling.main_func_var <- df.pert.effect.long %>%
  mutate(Perturbation = as.factor(Perturbation)) %>%
  mutate(Perturbation = fct_relevel(Perturbation, c("low", "high"))) %>%
  ggplot(aes(x = as.numeric(Perturbation), y = Value, color = Treatment)) +
  geom_point(size = 2.5) +
  geom_line(linewidth = 1, aes(linetype = Significance)) +
  #geom_text(data = df.pvalues, aes(x = x, y = y, label = denote, color = Treatment), hjust = 0, show.legend = F) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  scale_x_continuous(breaks = c(1, 2), labels = c("low", "high")) +
  xlab("Perturbation") +
  ylab("Relative length of the first semi-axis") +
  theme(axis.text = element_text(color = "black"),
        panel.grid = element_blank())+
  scale_linetype_manual(values = c("dashed","solid"))
```

# Section 4: dimensionality analysis using the correlation approach (func_var_det used)

This analysis will use the same data as that with pca approach. Here as correlation do not need to include all the stability components at the same time, so It does require to take out one functional variability component.

**4.1 compute the correlation coefficient** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
load(file = paste0(data.path, "/df.permutation.random.plots.Rdata"))
# calculate correlation coefficient
df.permutation %<>% mutate(corr.v = map(data, cor))
df.correlation <- df.permutation %>% select(-data)
```

**4.2 figure for correlation excluding func_var** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=8, fig.height=8}
# change the correlation matrix to long form and only keep three correlations
df.correlation %<>% mutate(corr.v2 = map(corr.v, function(m) {
  out <- as.data.frame(as.table(m))
  return(out)
}))
df.correlation %<>% select(-corr.v) %>% unnest(corr.v2)

df.correlation <- df.correlation %>%
  mutate(sta_pair = paste(Var1, Var2, sep = " & ")) %>%
  filter(sta_pair %in% c(
    "func_var & comp_var",
    "func_change & comp_change",
    "func_var & func_change",
    "comp_var & comp_change",
    "func_var_det & comp_var",
    "func_var_det & func_change",
    "func_change & comp_var",
    "func_var & comp_change",
    "func_var_det & comp_change"
  )) %>%
  select(-Var1, -Var2) %>%
  rename(Value = Freq) %>%
  mutate(Value.abs = abs(Value))

# factorize sta_pair
df.correlation <- df.correlation %>%
  mutate(sta_pair = str_replace_all(sta_pair, " & ", " &\n ")) %>%
  mutate(sta_pair = as.factor(sta_pair)) %>%
  mutate(sta_pair = fct_relevel(sta_pair, c(
    "func_var &\n comp_var",
    "func_change &\n comp_change",
    "func_var &\n func_change",
    "comp_var &\n comp_change",
    "func_var_det &\n comp_var",
    "func_var_det &\n func_change",
    "func_change &\n comp_var",
    "func_var &\n comp_change",
    "func_var_det &\n comp_change"
  )))

# a figure that showing the original change correlation coefficient (abs), excluding the ones involving func_var

corrs.all <- df.correlation$sta_pair %>% unique()
corrs.selected <- setdiff(corrs.all, corrs.all[str_detect(corrs.all, "func_var ")])

p.corr.change <- df.correlation %>%
  filter(final.year.set == 9, perm.id == 0, sta_pair %in% corrs.selected) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("N", "P", "K"))) %>%
  ggplot(aes(x = end_year, y = Value.abs, color = trt.set, linetype = PI.group)) +
  facet_wrap(~sta_pair, nrow = 1) +
  geom_line(linewidth = 1) +
  # geom_point(aes(shape = significance), size = 1.5) +
  scale_color_brewer(
    name = "Treatment:",
    palette = "Dark2",
    labels = c("+N", "+P", "+K")
  ) +
  # scale_shape_manual(values = c(4, 19)) +
  theme_bw() +
  theme(
    legend.position = "top", legend.box = "horizontal", legend.margin = margin(),
    legend.spacing.x = unit(1, "cm"), axis.text = element_text(color = "black", size = 10),
        panel.grid = element_blank()
  ) +
  xlab("Years") +
  ylab("Correlation strength") +
  labs(linetype = "Perturbation:") +
  guides(
    color = guide_legend(
      nrow =
        1
    ), # Arrange color legend in 1 row
    linetpye = guide_legend(nrow = 1) # Arrange shape legend in 1 row
  )
pdf(paste0(getwd(), "/myfigure/p.corr.chagne_func_var_det.pdf"), width = 8, height = 3)
p.corr.change
dev.off()
p.corr.change
p.corr.chagne_func_var_det <- p.corr.change

# nest df.correlation
df.nested <- df.correlation %>%
  select(-final.year.set) %>%
  group_by(trt.set, sta_pair, data.group, perm.id) %>%
  nest()

# perform the gls
# as some of the elements in df.nested will return errors, I here run GLS as a loop
# I also just extract the coefficient in the loop

low.set <- high.set <- results <- numeric(nrow(df.nested))

for (i in 1:nrow(df.nested)) {
  results[i] <- NA
  low.set[i] <- NA
  high.set[i] <- NA
  if (i %% 1000 == 0) print(i)
  try({
    dt_ <- df.nested$data[[i]]
    gls_model <- gls(Value.abs ~ PI.group,
      data = dt_,
      correlation = corAR1(form = ~ end_year | PI.group)
    )

    high.set[i] <- coef(gls_model)["(Intercept)"]
    results[i] <- coef(gls_model)["PI.groupLow"] # diff
    low.set[i] <- high.set[i] + results[i]
  })
}

df.nested$diff_value <- results
df.nested$low <- low.set
df.nested$high <- high.set

# note that the difference now is calculated as low-perturbation minus high-perturbation,
# I want to use it in an opposite way
df.nested <- df.nested %>%
  mutate(diff_value = -1 * diff_value)
# now factorize trt.set
df.nested <- df.nested %>%
  mutate(trt.set = paste0("+", trt.set)) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("+N", "+P", "+K")))

# only keep the key columns
df.results <- df.nested %>%
  ungroup() %>%
  # select(-data) %>%
  na.omit() %>% # remove the problametic gls result; only one
  group_by(trt.set, sta_pair) %>%
  nest()
# now add a column to represent the significance
df.results <- df.results %>%
  mutate(alpha = map_dbl(data, function(dt_) {
    x <- dt_ %>%
      filter(perm.id != 0) %>%
      select(diff_value) %>%
      unlist()
    x0 <- dt_ %>%
      filter(perm.id == 0) %>%
      select(diff_value) %>%
      unlist()
    length(which(x <= x0)) / length(x)
  }))
# also add a denote colomn for p values so I can add it to the figure later
df.results <- df.results %>%
  mutate(denote = map_chr(alpha, function(x) {
    x <- round(x, digits = 3)
    paste("P =", x)
  }))

df.results <- df.results %>% select(-data) # here

# now make a figure for null model result excluding the sta_pair involving func_var

# add a column represent significance groups
df.results <- df.results %>%
  mutate(sig.group = map_chr(alpha, function(x) {
    if (x < 0.05) {
      y <- "negative"
    } else if (x < 0.95) {
      y <- "non-significant"
    } else {
      y <- "positive"
    }
  }))

p.null <- df.nested %>%
  filter(data.group == "permutation", sta_pair %in% corrs.selected) %>%
  ggplot(aes(x = diff_value)) +
  facet_grid(trt.set ~ sta_pair) +
  geom_histogram(bins = 30, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% filter(data.group == "obs", sta_pair %in% corrs.selected) %>%
      left_join(df.results %>% select(trt.set, sta_pair, sig.group)),
    aes(xintercept = diff_value, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% mutate(x = -0.4, y = 140) %>% filter(sta_pair %in% corrs.selected),
    aes(x = x, y = y, label = denote),
    hjust = 0
  ) +
  theme_bw() +
  xlab("Difference in correlation strength") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "bottom",
        panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black", "red")) +
  labs(color = "Significance:")

# combine the null distribution result with that of the change in correlation along time and perturbation strength
p.corr.stats_func_var_det <- p.null

SI.correlation_func_var_det <- p.corr.chagne_func_var_det / p.corr.stats_func_var_det +
  plot_layout(heights = c(1, 2)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(face = "bold"))

pdf(paste0(getwd(), "/myfigure/SI.correlation_func_var_det.pdf"), width = 8, height = 8)
SI.correlation_func_var_det
dev.off()
SI.correlation_func_var_det
```

Now a figure for the main effect of perturbation on correlation, which will be later combined with the main effect of perturbaiton on overall decoupling.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=6, fig.height=6}
df.pert.effect <- df.nested %>%
  filter(data.group == "obs") %>%
  ungroup()
df.pert.effect.long <- df.pert.effect %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, sta_pair, high, low) %>%
  pivot_longer(cols = c(high, low), names_to = "Perturbation", values_to = "Value.abs")

df.pvalues <- df.results %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, sta_pair, denote) %>%
  arrange(desc(Treatment))
df.pvalues$x <- rep(1.6, 27)
df.pvalues$y <- rep(c(0.3, 0.45, 0.6), each = 9)

df.pert.effect.long <- df.pert.effect.long %>% 
  left_join(df.results %>%
              rename(Treatment = trt.set) %>% select(sta_pair, Treatment, alpha))
df.pert.effect.long <- df.pert.effect.long %>% 
  mutate(Significance = map_chr(alpha, function(x)ifelse(x<0.05|x>0.95, "significant","non-significant")))

P.corr.main_func_var_det <- df.pert.effect.long %>%
  filter(sta_pair %in% corrs.selected) %>%
  mutate(Perturbation = as.factor(Perturbation)) %>%
  mutate(Perturbation = fct_relevel(Perturbation, c("low", "high"))) %>%
  ggplot(aes(x = as.numeric(Perturbation), y = Value.abs, color = Treatment)) +
  geom_line(linewidth = 1, aes(linetype = Significance)) +
  geom_point(size = 2.5) +
  facet_wrap(~sta_pair, nrow = 3) +
  #geom_text(
  #  data = df.pvalues %>% filter(sta_pair %in% corrs.selected),
   # aes(x = x, y = y, label = denote, color = Treatment), hjust = 0, show.legend = F, size = 3
 # ) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  scale_x_continuous(breaks = c(1, 2), labels = c("low", "high")) +
  xlab("Perturbation") +
  ylab("Correlation strength") +
  theme(
    axis.text = element_text(color = "black"),
    panel.spacing.x = unit(.5, "cm"),
        panel.grid = element_blank()
  )+
  scale_linetype_manual(values = c("dashed","solid"))+
  ylim(c(0,0.65))
# now combine with the main figure on decoupling using semi-axis approach

# need to solve the collect guide problem or use cowplot to add a legend
# collect guides not working for these ggplots
# so I make another ggplot and extract legend and add the legend to the composite figure
Main.decoupling_func_var_det <- P.decoupling.main_func_var_det +
  P.corr.main_func_var_det +
  plot_annotation(tag_levels = "a") +
  plot_layout(widths = c(1, 2), guides = "collect") & # Collect legends into one
  theme(plot.tag = element_text(face = "bold"), legend.position = "none")
new.figure <- df.pert.effect.long %>%
  filter(sta_pair %in% corrs.selected) %>%
  mutate(Perturbation = as.factor(Perturbation)) %>%
  mutate(Perturbation = fct_relevel(Perturbation, c("low", "high"))) %>%
  ggplot(aes(x = as.numeric(Perturbation), y = Value.abs, color = Treatment)) +
  geom_line(aes(linetype = Significance)) +
  geom_point() +
  facet_wrap(~sta_pair) +
  theme_bw() +
  theme(legend.position = "top") +
  scale_color_brewer(palette = "Dark2")+
  scale_linetype_manual(values = c("dashed","solid"))

p.legend <- get_plot_component(new.figure, "guide-box-top", return_all = TRUE)
pdf(paste0(getwd(), "/myfigure/Main.decoupling_func_var_det.pdf"), width = 6, height = 6)
plot_grid(Main.decoupling_func_var_det, p.legend, ncol = 1, rel_heights = c(10, 1))
dev.off()
plot_grid(Main.decoupling_func_var_det, p.legend, ncol = 1, rel_heights = c(10, 1))
```


**4.3 figure for correlation involving func_var** 

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
# a figure that showing the original change correlation coefficient (abs), the ones involving func_var

corrs.selected2 <- setdiff(corrs.all, corrs.selected)

p.corr.change <- df.correlation %>%
  filter(final.year.set == 9, perm.id == 0, sta_pair %in% corrs.selected2) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("N", "P", "K"))) %>%
  ggplot(aes(x = end_year, y = Value.abs, color = trt.set, linetype = PI.group)) +
  facet_wrap(~sta_pair, nrow = 1) +
  geom_line(linewidth = 1) +
  # geom_point(aes(shape = significance), size = 1.5) +
  scale_color_brewer(
    name = "Treatment:",
    palette = "Dark2",
    labels = c("+N", "+P", "+K")
  ) +
  # scale_shape_manual(values = c(4, 19)) +
  theme_bw() +
  theme(
    legend.position = "top", legend.box = "horizontal", legend.margin = margin(), legend.spacing.x = unit(1, "cm"),
    axis.text = element_text(color = "black", size = 10),
    panel.spacing.x = unit(.25, "cm"),
        panel.grid = element_blank()
  ) +
  xlab("Years") +
  ylab("Correlation strength") +
  labs(linetype = "Perturbation:") +
  guides(
    color = guide_legend(
      nrow =
        1
    ), # Arrange color legend in 1 row
    linetpye = guide_legend(nrow = 1) # Arrange shape legend in 1 row
  )
pdf(paste0(getwd(), "/myfigure/p.corr.chagne_func_var.pdf"), width = 8, height = 3)
p.corr.change
dev.off()
p.corr.change
p.corr.chagne_func_var <- p.corr.change


p.null <- df.nested %>%
  filter(data.group == "permutation", sta_pair %in% corrs.selected2) %>%
  ggplot(aes(x = diff_value)) +
  facet_grid(trt.set ~ sta_pair) +
  geom_histogram(bins = 30, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% filter(data.group == "obs", sta_pair %in% corrs.selected2) %>%
      left_join(df.results %>% select(trt.set, sta_pair, sig.group)),
    aes(xintercept = diff_value, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% mutate(x = -0.47, y = 100) %>% filter(sta_pair %in% corrs.selected2),
    aes(x = x, y = y, label = denote),
    hjust = 0
  ) +
  theme_bw() +
  xlab("Difference in correlation strength") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "bottom",
    panel.spacing.x = unit(.25, "cm"),
        panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black", "red")) +
  labs(color = "Significance:") +
  scale_x_continuous(breaks = c(-0.4, 0, 0.4))

# combine the null distribution result with that of the change in correlation along time and perturbation strength
p.corr.stats_func_var <- p.null

SI.correlation_func_var <- p.corr.chagne_func_var / p.corr.stats_func_var +
  plot_layout(heights = c(1, 2)) +
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(face = "bold"))

pdf(paste0(getwd(), "/myfigure/SI.correlation_func_var.pdf"), width = 7, height = 7)
SI.correlation_func_var
dev.off()
SI.correlation_func_var
```

Now a figure for the main effect of perturbation on correlation, which will be later combined with the main effect of perturbaiton on overall decoupling.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
df.pert.effect <- df.nested %>%
  filter(data.group == "obs") %>%
  ungroup()
df.pert.effect.long <- df.pert.effect %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, sta_pair, high, low) %>%
  pivot_longer(cols = c(high, low), names_to = "Perturbation", values_to = "Value.abs")

df.pvalues <- df.results %>%
  rename(Treatment = trt.set) %>%
  select(Treatment, sta_pair, denote) %>%
  arrange(desc(Treatment))
df.pvalues$x <- rep(1.6, 27)
df.pvalues$y <- rep(c(0.3, 0.45, 0.6), each = 9)

df.pert.effect.long <- df.pert.effect.long %>% 
  left_join(df.results %>%
              rename(Treatment = trt.set) %>% select(sta_pair, Treatment, alpha))
df.pert.effect.long <- df.pert.effect.long %>% 
  mutate(Significance = map_chr(alpha, function(x)ifelse(x<0.05|x>0.95, "significant","non-significant")))

P.corr.main_func_var <- df.pert.effect.long %>%
  filter(sta_pair %in% corrs.selected2) %>%
  mutate(Perturbation = as.factor(Perturbation)) %>%
  mutate(Perturbation = fct_relevel(Perturbation, c("low", "high"))) %>%
  ggplot(aes(x = as.numeric(Perturbation), y = Value.abs, color = Treatment)) +
  geom_line(linewidth = 1, aes(linetype = Significance)) +
  geom_point(size = 2.5) +
  facet_wrap(~sta_pair, nrow = 3) +
  #geom_text(
  #  data = df.pvalues %>% filter(sta_pair %in% corrs.selected2),
  #  aes(x = x, y = y, label = denote, color = Treatment), hjust = 0, show.legend = F, size = 3
  #) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  scale_x_continuous(breaks = c(1, 2), labels = c("low", "high")) +
  xlab("Perturbation") +
  ylab("Correlation strength") +
  theme(
    axis.text = element_text(color = "black"),
    panel.spacing.x = unit(.5, "cm"),
        panel.grid = element_blank()
  )+
  scale_linetype_manual(values = c("dashed","solid"))
# now combine with the main figure on decoupling using semi-axis approach

Main.decoupling_func_var <- P.decoupling.main_func_var +
  P.corr.main_func_var +
  plot_annotation(tag_levels = "a") +
  plot_layout(widths = c(1, 1), guides = "collect") & # Collect legends into one
  theme(plot.tag = element_text(face = "bold"), legend.position = "none")

pdf(paste0(getwd(), "/myfigure/Main.decoupling_func_var.pdf"), width = 6, height = 6)
plot_grid(Main.decoupling_func_var, p.legend, ncol = 1, rel_heights = c(10, 1))
dev.off()
plot_grid(Main.decoupling_func_var, p.legend, ncol = 1, rel_heights = c(10, 1))
```
