---
title: "R codes for the NutNet-Stability project"
subtitle: "Step4.3: Test effect of time on complexity of ecological stability"
author: "Qiang Yang"
date: "`r Sys.Date()`"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{}
- \fancyfoot[CO,CE]{R codes for the NutNet-stability project}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

In the main manuscript, we only presented the results of perturbation intensity on the complexity (dimensionality) of ecological stability. We need to show why time is not considered as a main effect. In this Rmd file, I will conduct analysis for this purpose.


# Section 1: setup work environment

```{r, warning=FALSE, message=FALSE, highlight=T, eval=T}
rm(list = ls())
gc()
# necessary R packages
Packages <- c(
  "tidyverse", "magrittr", "grid", "gridExtra", "scales",
  "broom", "dlookr", "nlme", "patchwork", "cowplot"
)
pacman::p_load(char = Packages)
# functions with conflicting status (presence in different packages used here)
select <- dplyr::select
summarise <- dplyr::summarise
# working path
main.path <- getwd()
data.path <- paste0(main.path, "/mydata")
figure.path <- paste0(main.path, "/myfigure")
set.seed(1234)
```



# Section 2: dimensionality analysis using the PCA approach (func_var_det used)

There are two functional variability. Here I will use func_var_det. But I will use func_var in section 3 to check if conclusion will be changed.

**2.1 data organization - observed data**

Note that all stability components here are first scaled so that they have the similar scale.
```{r, warning=FALSE, message=FALSE, eval=F, highlight=T, fig.width=7, fig.height=4}
load(paste0(data.path, "/df.tidy.stabilities.Rdata"))
load(paste0(data.path, "/df.tidy.Rdata"))
df.full <- df.tidy.stabilities %>%
  left_join(df.tidy %>% select(-contains("data")))
rm(df.tidy, df.tidy.stabilities)

df.full %<>%
  select(site_code:end_year, matches("_added_proportion|change|var"))
sta.names <- colnames(df.full) %>% str_subset("change|var")

# For the PCA analysis, all stability components need to be available for each row
# no NA, no infinite
# remove the relevant cases
df.full <- df.full %>%
  filter_all(all_vars(!is.infinite(.))) %>%
  filter_all(all_vars(!is.na(.)))

df.stability <- df.full

# scale stability components
df.stability %<>% mutate(across(one_of(sta.names), scale))


# make a meta data for multidimensionality analysis
trt.set <- c("N", "P", "K")
final.year.set <- 9 # original using 5:9
df.meta <- crossing(trt.set, final.year.set)

## add data, nested by year

df.meta <- df.meta %>%
  mutate(data = map2(
    trt.set, final.year.set,
    function(x1, x2) {
      plots.to.keep <- df.stability %>%
        filter(
          trt == x1,
          end_year == x2
        ) %>%
        select(site_code, block, trt, plot)
      subdat <- df.stability %>%
        semi_join(plots.to.keep) %>%
        filter(
          trt == x1,
          end_year <= x2
        )
      return(subdat)
    }
  ))
df.meta %<>% ungroup()

## add data, nested by perturbation intensity and end_year

df.meta <- df.meta %>%
  mutate(data2 = map2(
    trt.set, data,
    function(x1, dat) {
      df.pert <- dat %>%
        select(site_code, block, trt, plot, paste0(x1, "_added_proportion")) %>%
        distinct() %>%
        set_colnames(c("site_code", "block", "trt", "plot", "PI"))
      df.pert %<>%
        mutate(PI.group = ifelse(PI > median(PI), "High", "Low"))
      dat %<>% left_join(df.pert)
      dat.nested <- dat %>%
        group_by(end_year, PI.group) %>%
        nest() %>%
        ungroup()
      return(dat.nested)
    }
  ))

df.meta %<>%
  select(-data) %>%
  unnest(cols = data2)

df.N.P.K <- df.meta %>% ungroup()

# # further only select stability components from data
df.N.P.K %<>%
  mutate(data = map(data, function(dat) dat %>% select(one_of(sta.names))))
df.N.P.K.obs <- df.N.P.K
```

**2.2 data organization - data produced by permutation**

Conduct permutations. The dataset full were permuted by plot*final.year.set*end_year.

```{r, warning=FALSE, message=FALSE, eval=F, highlight=T, fig.width=7, fig.height=4}
set.seed(1234)
L <- list()
for (i in 1:1000) { # about 15 minutes to finish
  # Specify the column names to permute
  print(i)
  print(Sys.time())
  df.stability.obs <- df.stability
# first randomize the nutrient perturbaiton intensity of plots
df.nutrient.pert <- df.stability.obs %>% 
  select(site_code, block, trt, plot, contains("proportion")) %>% 
  distinct()# the nutrient perturbation of each plot
df.nutrient.pert.random <- df.nutrient.pert %>% 
  group_by(trt) %>% 
  nest() %>% 
  mutate(data2 = map(data, function(dt_){
    dt_ %>% 
      mutate(N_added_proportion = sample(N_added_proportion),
             P_added_proportion = sample(P_added_proportion),
             K_added_proportion = sample(K_added_proportion))
  })) %>% 
  select(-data) %>% 
  unnest(cols = data2)
df.stability.random.plots <- df.stability.obs %>% 
  select(-contains("proportion")) %>% 
  left_join(df.nutrient.pert.random)

# then randomize years within each plot
# here I want the configuration of the stability components unchanged
# so only need to randomize the year column within each plot

# my initial codes have some problem with the situation when the plot only has one end_year
# so I took a two-step approach
df.stability.onlyone.year <- df.stability.random.plots %>% 
  count(site_code, block, trt, plot) %>% 
  filter(n == 1) %>% 
  left_join(df.stability.random.plots) %>% 
  select(-n)

df.stability.random.plots.AND.years <- df.stability.random.plots%>% 
  anti_join(df.stability.onlyone.year) %>% 
  group_by(site_code, block, trt, plot) %>% 
  nest() %>% 
  mutate(data = map(data, function(dt_){
    dt_ %>% mutate(end_year = sample(end_year))
  })) %>% 
  unnest(cols = data)

df.stability.0 <- df.stability.onlyone.year %>% bind_rows(df.stability.random.plots.AND.years)

  # make a meta data for multidimensionality analysis
  trt.set <- c("N", "P", "K")
  final.year.set <- 9 # original using 5:9
  df.meta <- crossing(trt.set, final.year.set)

  ## add data, nested by year

  df.meta <- df.meta %>%
    mutate(data = map2(
      trt.set, final.year.set,
      function(x1, x2) {
        plots.to.keep <- df.stability.0 %>%
          filter(
            trt == x1,
            end_year == x2
          ) %>%
          select(site_code, block, trt, plot)
        subdat <- df.stability.0 %>%
          semi_join(plots.to.keep) %>%
          filter(
            trt == x1,
            end_year <= x2
          )
        return(subdat)
      }
    ))
  df.meta %<>% ungroup()

  ## add data, nested by perturbation intensity and end_year

  df.meta <- df.meta %>%
    mutate(data2 = map2(
      trt.set, data,
      function(x1, dat) {
        df.pert <- dat %>%
          select(site_code, block, trt, plot, paste0(x1, "_added_proportion")) %>%
          distinct() %>%
          set_colnames(c("site_code", "block", "trt", "plot", "PI"))
        df.pert %<>%
          mutate(PI.group = ifelse(PI > median(PI), "High", "Low"))
        dat %<>% left_join(df.pert)
        dat.nested <- dat %>%
          group_by(end_year, PI.group) %>%
          nest() %>%
          ungroup()
        return(dat.nested)
      }
    ))

  df.meta %<>%
    select(-data) %>%
    unnest(cols = data2)

  df.N.P.K <- df.meta %>% ungroup()

  # # further only select stability components from data
  df.N.P.K %<>%
    mutate(data = map(data, function(dat) dat %>% select(one_of(sta.names))))

  L[[i]] <- df.N.P.K
  rm(df.N.P.K)
  rm(df.stability.0)
}

names(L) <- 1:1000
df.N.P.K.perm <- tibble(
  perm.id = names(L), # This will be the ID column
  data = L # This will be the column containing data.frames
) %>%
  unnest(data) %>%
  mutate(perm.id = as.numeric(perm.id))

df.permutation <- df.N.P.K.obs %>%
  mutate(
    data.group = "obs",
    perm.id = 0
  ) %>%
  bind_rows(df.N.P.K.perm %>%
    mutate(data.group = "permutation"))
save(df.permutation, file = paste0(data.path, "/df.permutation.random.plots.AND.years.Rdata"))
```

**2.3 compute the covariance matrix and the eigen vector of the covariance matrix (func_var_det)** 

Here I will choose func_var_det.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
load(file = paste0(data.path, "/df.permutation.random.plots.AND.years.Rdata"))
df.permutation <- df.permutation %>%
  mutate(data = map(data, function(dt1) dt1 %>% select(-func_var)))
# calculate covariance matrix and eigenvector
df.permutation %<>% mutate(cov.M = map(data, cov))
df.permutation %<>% mutate(eigvec = map(cov.M, function(x) eigen(x)$vectors))
#### calculate the variance explained by first eigen vector of the covariance matrix
df.permutation %<>%
  mutate(var.explained = map2_dbl(data, eigvec, function(D, X) {
    D <- as.matrix(D)
    P <- D %*% X
    V <- apply(P, 2, var)
    V[1] / sum(V)
  }))

df.variance.explained <- df.permutation %>%
  select(trt.set, final.year.set, end_year, PI.group, data.group, perm.id, var.explained)
```

**2.4 a simple regression (func_var_det)** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=5}
df.nested <- df.variance.explained %>%
  select(-final.year.set) %>% # unique value, so can be removed
  mutate(PI.group = as.factor(PI.group),
         PI.group = fct_relevel(PI.group, c("Low", "High"))) %>% 
  group_by(trt.set, perm.id, data.group) %>%
  nest() 
# conduct regression
df.nested <- df.nested %>% 
  mutate(model.set = map(data, function(.dt){
    .dt <- .dt %>% mutate(end_year = scale(end_year))
    lm(data=.dt, var.explained~end_year*PI.group)
  })) %>% 
  mutate(coef.set = map(model.set, tidy)) %>% 
  select(-model.set, -data) %>% 
  unnest(cols = coef.set)


# now factorize trt.set
df.nested <- df.nested %>%
  mutate(trt.set = paste0("+", trt.set)) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("+N", "+P", "+K")))

# only keep the key columns
df.results <- df.nested %>%
  group_by(trt.set, term) %>%
  nest()
# now add a column to represent the significance
df.results <- df.results %>%
  mutate(alpha = map_dbl(data, function(dt_) {
    x <- dt_ %>%
      filter(perm.id != 0) %>%
      select(estimate) %>%
      unlist()
    x0 <- dt_ %>%
      filter(perm.id == 0) %>%
      select(estimate) %>%
      unlist()
    length(which(x <= x0)) / length(x)
  }))
# also add a denote colomn for p values so I can add it to the figure later
df.results <- df.results %>%
  mutate(denote = map_chr(alpha, function(x) {
    paste("P =", x)
  }))
# add a column represent significance groups
df.results <- df.results %>%
  mutate(sig.group = map_chr(alpha, function(x) {
    if (x < 0.05) {
      y <- "negative"
    } else if (x < 0.95) {
      y <- "non-significant"
    } else {
      y <- "positive"
    }
  }))

# now make a figure
predictor.labs <- c("Year", "High Perturbation", "Year:High_perturbation")
names(predictor.labs) <- c("end_year", "PI.groupHigh", "end_year:PI.groupHigh")

p.null <- df.nested %>%
  filter(data.group == "permutation", term!="(Intercept)") %>%
  ggplot(aes(x = estimate)) +
  facet_grid(trt.set~term,labeller = labeller(term = predictor.labs)) +
  geom_histogram(bins = 50, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% 
      filter(data.group == "obs", term!="(Intercept)") %>% 
      left_join(df.results %>% filter(term!="(Intercept)")%>% select(trt.set, sig.group)),
    aes(xintercept = estimate, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% filter(term!="(Intercept)") %>% mutate(x = -0.1, y = 150),
    aes(x = x, y = y, label = denote)
  ) +
  theme_bw() +
  xlab("Model coefficient estimate") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "top",
    panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black")) +
  labs(color = "Significance:")

pdf(paste0(getwd(), "/myfigure/preliminary.decoupling.analysis_func_var_det.pdf"), width = 7, height = 5)
p.null
dev.off()
p.null
```


**2.5 compute the covariance matrix and the eigen vector of the covariance matrix (func_var)** 

Here I will choose func_var_det.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=4}
load(file = paste0(data.path, "/df.permutation.random.plots.AND.years.Rdata"))
df.permutation <- df.permutation %>%
  mutate(data = map(data, function(dt1) dt1 %>% select(-func_var_det)))
# calculate covariance matrix and eigenvector
df.permutation %<>% mutate(cov.M = map(data, cov))
df.permutation %<>% mutate(eigvec = map(cov.M, function(x) eigen(x)$vectors))
#### calculate the variance explained by first eigen vector of the covariance matrix
df.permutation %<>%
  mutate(var.explained = map2_dbl(data, eigvec, function(D, X) {
    D <- as.matrix(D)
    P <- D %*% X
    V <- apply(P, 2, var)
    V[1] / sum(V)
  }))

df.variance.explained <- df.permutation %>%
  select(trt.set, final.year.set, end_year, PI.group, data.group, perm.id, var.explained)
```

**2.6 a simple regression (func_var)** 
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, fig.width=7, fig.height=5}
df.nested <- df.variance.explained %>%
  select(-final.year.set) %>% # unique value, so can be removed
  mutate(PI.group = as.factor(PI.group),
         PI.group = fct_relevel(PI.group, c("Low", "High"))) %>% 
  group_by(trt.set, perm.id, data.group) %>%
  nest() 
# conduct regression
df.nested <- df.nested %>% 
  mutate(model.set = map(data, function(.dt){
    .dt <- .dt %>% mutate(end_year = scale(end_year))
    lm(data=.dt, var.explained~end_year*PI.group)
  })) %>% 
  mutate(coef.set = map(model.set, tidy)) %>% 
  select(-model.set, -data) %>% 
  unnest(cols = coef.set)


# now factorize trt.set
df.nested <- df.nested %>%
  mutate(trt.set = paste0("+", trt.set)) %>%
  mutate(trt.set = as.factor(trt.set)) %>%
  mutate(trt.set = fct_relevel(trt.set, c("+N", "+P", "+K")))

# only keep the key columns
df.results <- df.nested %>%
  group_by(trt.set, term) %>%
  nest()
# now add a column to represent the significance
df.results <- df.results %>%
  mutate(alpha = map_dbl(data, function(dt_) {
    x <- dt_ %>%
      filter(perm.id != 0) %>%
      select(estimate) %>%
      unlist()
    x0 <- dt_ %>%
      filter(perm.id == 0) %>%
      select(estimate) %>%
      unlist()
    length(which(x <= x0)) / length(x)
  }))
# also add a denote colomn for p values so I can add it to the figure later
df.results <- df.results %>%
  mutate(denote = map_chr(alpha, function(x) {
    paste("P =", x)
  }))
# add a column represent significance groups
df.results <- df.results %>%
  mutate(sig.group = map_chr(alpha, function(x) {
    if (x < 0.05) {
      y <- "negative"
    } else if (x < 0.95) {
      y <- "non-significant"
    } else {
      y <- "positive"
    }
  }))

# now make a figure
predictor.labs <- c("Year", "High Perturbation", "Year:High_perturbation")
names(predictor.labs) <- c("end_year", "PI.groupHigh", "end_year:PI.groupHigh")

p.null <- df.nested %>%
  filter(data.group == "permutation", term!="(Intercept)") %>%
  ggplot(aes(x = estimate)) +
  facet_grid(trt.set~term,labeller = labeller(term = predictor.labs)) +
  geom_histogram(bins = 50, fill = "lightgray", color = "black") +
  geom_vline(
    data = df.nested %>% 
      filter(data.group == "obs", term!="(Intercept)") %>% 
      left_join(df.results %>% filter(term!="(Intercept)")%>% select(trt.set, sig.group)),
    aes(xintercept = estimate, color = sig.group),
    linewidth = 1
  ) +
  geom_text(
    data = df.results %>% filter(term!="(Intercept)") %>% mutate(x = -0.1, y = 150),
    aes(x = x, y = y, label = denote)
  ) +
  theme_bw() +
  xlab("Model coefficient estimate") +
  ylab("Count") +
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "top",
    panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("blue", "black", "red")) +
  labs(color = "Significance:")

pdf(paste0(getwd(), "/myfigure/preliminary.decoupling.analysis_func_var.pdf"), width = 7, height = 5)
p.null
dev.off()
p.null
```