---
title: "R codes for the NutNet-Stability project"
subtitle: "Step2: Quantify stability components"
author: "Qiang Yang"
date: "`r Sys.Date()`"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{}
- \fancyfoot[CO,CE]{R codes for the NutNet-stability project - quantify stability components}
- \fancyfoot[LE,RO]{\thepage}
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage



# Setup work environment

```{r, warning=FALSE, message=FALSE, highlight=T, prompt=TRUE, eval=T}
rm(list = ls())
gc()
# necessary R packages
Packages <- c("tidyverse", "magrittr", "vegan", "dlookr")
pacman::p_load(char = Packages)
# functions with conflicting status (presence in different packages used here)
select <- dplyr::select
summarise <- dplyr::summarise
# working path
main.path <- getwd()
data.path <- paste0(main.path, "/mydata")
set.seed(1234)
```

\newpage


# Section 1: Quantify stability components using biomass data

This include the functional variability (detrended and non-detrended) and functional resistance.

**1.1 temporal functional variability**

Temporal functional variability was quantified as: CV.treatment / CV.control using total biomass.

There might be multiple control plots in one block. First, select one control for each block.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
load(file = paste0(data.path, "/df.tidy.Rdata"))
df.tidy.mass <- df.tidy %>% #only the plots with biomass data
  filter(data.availability != "only coverage data") %>% 
  select(site_code:trt,biomass.data,total.mass) %>% 
  rename(data = biomass.data)
df.controls <- df.tidy.mass %>%
  unnest(cols=data) %>% 
  ungroup() %>% 
  filter(trt == "Control") %>%
  select(site_code, block, plot, year_trt) %>%
  distinct()
set.seed(1234)
df.controls.selected <- df.controls %>% # select the plot with largest year_trt
  arrange(site_code, block, desc(year_trt)) %>%
  group_by(site_code, block) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(-year_trt)
```

Then calculate the coefficient of variation in each plot, both originally and detrendedly.
```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.sta.variability <- df.tidy.mass %>%
  select(site_code, block, plot, trt, data) %>%
  mutate(results = map(data, function(dat) {
    df <- dat %>%
      group_by(year_trt) %>%
      summarise(total.mass = sum(mass)) %>%
      ungroup() # calculate the total mass at each year
    output <- df %>%
      select(year_trt) %>%
      rename(end_year = year_trt) %>%
      filter(end_year >= 4) %>%
      mutate(VAR.original = map_dbl(end_year, function(ey) {
        df.sub <- df %>%
          filter(year_trt >= 1, year_trt <= ey)
        y <- df.sub$total.mass
        sd(y) / mean(y)
      })) %>%
      mutate(VAR.detrended = map_dbl(end_year, function(ey) {
        df.sub <- df %>%
          filter(year_trt >= 1, year_trt <= ey)
        x <- df.sub$year_trt
        y <- df.sub$total.mass
        lmf <- lm(y ~ x)
        residuals <- lmf$residuals
        sd(residuals) / mean(y)
      }))
  }))
df.sta.variability %<>%
  select(-data) %>%
  unnest(cols = results)

df.sta.variability.treatment <- df.sta.variability %>%
  filter(trt %in% c("N", "P", "K"))
df.sta.variability.control <- df.sta.variability %>%
  left_join(df.controls.selected, .)

df.sta.variability.control %<>%
  select(-trt, -plot) %>%
  rename(
    VAR.original.control = VAR.original,
    VAR.detrended.control = VAR.detrended
  )

df.sta.variability.treatment %<>%
  left_join(df.sta.variability.control)
df.sta.variability.treatment %<>% ungroup()
df.sta.variability.treatment %<>% mutate(
  temporal.functional.variability = VAR.original / VAR.original.control,
  temporal.functional.variability.detrended = VAR.detrended / VAR.detrended.control
)

# it is possible that some plots of +N, +P, and +K have longer sampling years. 
# In these cases, the code will return a NA value for the stability quantified -
# using the time range including the year when the control has no data.
df.sta.temporal.functional.variability.mass <- df.sta.variability.treatment %>%
  na.omit()
```



**2.2 extent of functional change**

The extent of functional change was quantified as the mean of the log-response-ratio between the total mass of the treatment plot and the control plot acrossing survey years.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.tidy.mass.control <- df.tidy.mass %>% left_join(df.controls.selected, .)
df.tidy.mass.control %<>%
  select(site_code, block, plot, data) %>%
  rename(control.plot = plot, control.data = data)
df.tidy.mass.2 <- df.tidy.mass %>%
  left_join(df.tidy.mass.control) %>%
  select(site_code, block, plot, control.plot, trt, data, control.data)

df.mean.lrr <- df.tidy.mass.2 %>%
  filter(trt %in% c("N", "P", "K")) %>%
  mutate(results = pmap(list(trt, data, control.data), function(trt.v, dat1, dat2) {
    dat <- dat1 %>%
      mutate(g = "experimental") %>%
      bind_rows(dat2 %>% mutate(g = "control")) %>%
      select(year_trt, g, category, mass) %>%
      filter(year_trt %in% intersect(dat1$year_trt, dat2$year_trt))
    dat.nested <- dat %>%
      group_by(year_trt, g) %>%
      summarise(total_mass = sum(mass))
    lRR.total.mass <- dat.nested %>%
      pivot_wider(names_from = g, values_from = total_mass) %>%
      mutate(lrr = log(experimental / control))
    output <- lRR.total.mass %>%
      select(year_trt) %>%
      rename(end_year = year_trt) %>%
      filter(end_year >= 4) %>%
      mutate(mean.lrr = map_dbl(end_year, function(x) {
        lRR.total.mass %>%
          filter(year_trt <= x, year_trt >= 1) %>%
          pull(lrr) %>%
          abs() %>%
          mean()
      }))
    return(output)
  }))

df.sta.mean.lrr <- df.mean.lrr %>%
  select(-data, -control.plot, -control.data) %>%
  unnest(cols = results)

df.sta.extent.functional.change.mass <- df.sta.mean.lrr %>%
  rename(extent.functional.change = mean.lrr)

# if the biomass of some plots is zero, the returned stability values will be NA
df.sta.extent.functional.change.mass <- df.sta.extent.functional.change.mass %>%
  na.omit()
```

# Section 2: Quantify stability components using cover data

This includes the compositional variability (detrended and non-detrended) and compositional resistance.

**2.1 compositional variability**

As Turnover.treatment/Turnover.control.

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.tidy.coverage <- df.tidy %>% 
  filter(data.availability != "only biomass data") %>% 
  select(site_code:trt, coverage.data) %>% 
  rename(data = coverage.data)
df.controls <- df.tidy.coverage %>%
  unnest(cols=data) %>% 
  ungroup() %>% 
  filter(trt == "Control") %>%
  select(site_code, block, plot, year_trt) %>%
  distinct()
set.seed(1234)
df.controls.selected <- df.controls %>% # select the plot with largest year_trt
  arrange(site_code, block, desc(year_trt)) %>%
  group_by(site_code, block) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(-year_trt)
df.sta.compositional.turnover <- df.tidy.coverage %>%
  mutate(results = map(data, function(dat) {
    M <- dat %>% # long to wide
      select(year_trt, Taxon, max_cover) %>%
      pivot_wider(names_from = Taxon, values_from = max_cover, values_fill = 0)
    # jaccard distance between community at year_trt
    dist.m <- vegdist(M %>% select(-year_trt),
      method = "jaccard",
      binary = TRUE
    )
    dist.m %<>% as.matrix()
    # only dist between consecutive years
    dist.v <- diag(dist.m[2:nrow(dist.m), 1:(nrow(dist.m) - 1)])
    df <- M %>%
      select(year_trt) %>%
      filter(year_trt != 0) %>%
      arrange(year_trt) %>%
      mutate(dist = dist.v)
    output <- df %>%
      select(year_trt) %>%
      rename(end_year = year_trt) %>%
      filter(end_year >= 4) %>%
      mutate(compositional.turnover = map_dbl(end_year, function(ey) {
        df %>%
          filter(year_trt <= ey) %>%
          pull(dist) %>%
          mean()
      }))
  }))
df.sta.compositional.turnover %<>%
  select(-data) %>%
  unnest(cols = results)

df.sta.compositional.turnover.treatment <- df.sta.compositional.turnover %>%
  filter(trt %in% c("N", "P", "K"))
df.sta.compositional.turnover.control <- df.sta.compositional.turnover %>%
  semi_join(df.controls.selected)

df.sta.compositional.turnover.control %<>%
  select(-trt, -plot) %>%
  rename(compositional.turnover.control = compositional.turnover)

df.sta.compositional.turnover.treatment %<>%
  left_join(df.sta.compositional.turnover.control)
df.sta.compositional.turnover.treatment %<>%
  mutate(
    temporal.compositional.variability =
      compositional.turnover / compositional.turnover.control
  )

df.sta.temporal.compositional.variability <- df.sta.compositional.turnover.treatment
# as the survey range of the control plot is shorter than the treatment plots, - 
# there will be some NA values for the stability quantified using long range data
df.sta.temporal.compositional.variability <- na.omit(df.sta.temporal.compositional.variability)
```

**2.2 extent of compositional change**

the mean dissimilarity between the treatment plot and the control plot acrossing survey years; dissimilarity is using Bray-Curtis dissimilarity of log(cover+1)

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}

df.tidy.control <- df.tidy.coverage %>% semi_join(df.controls.selected)
df.tidy.control %<>%
  select(site_code, block, plot, data) %>%
  rename(control.plot = plot, control.data = data)
df.tidy.coverage.2 <- df.tidy.coverage %>%
  left_join(df.tidy.control) %>%
  select(site_code, block, plot, control.plot, trt, data, control.data)

df.max.mean.bc.numeric <- df.tidy.coverage.2 %>%
  filter(trt != "Control") %>%
  mutate(results = pmap(list(trt, data, control.data), function(trt.v, dat1, dat2) {
    dat <- dat1 %>%
      mutate(g = "experimental") %>%
      bind_rows(dat2 %>% mutate(g = "control")) %>%
      select(year_trt, g, Taxon, max_cover) %>%
      filter(year_trt %in% intersect(dat1$year_trt, dat2$year_trt))
    dat.wide <- dat %>%
      pivot_wider(names_from = "Taxon", values_from = "max_cover", values_fill = 0)
    dat.nested <- dat.wide %>%
      group_by(year_trt) %>%
      nest()
    dat.nested %<>%
      mutate(bc.dist = map_dbl(data, function(x) {
        x <- x %>%
          select(-g) %>%
          as.matrix()
        x <- log(x + 1)
        out <- x %>%
          vegdist(., method = "bray", binary = F, diag = TRUE, upper = TRUE) %>%
          as.matrix()
        out[1, 2]
      }))
    output <- dat.nested %>%
      select(year_trt) %>%
      rename(end_year = year_trt) %>%
      filter(end_year >= 4) %>%
      mutate(max.bc.numeric = map_dbl(end_year, function(x) {
        dat.nested %>%
          filter(year_trt <= x, year_trt >= 1) %>%
          pull(bc.dist) %>%
          max()
      })) %>%
      mutate(mean.bc.numeric = map_dbl(end_year, function(x) {
        dat.nested %>%
          filter(year_trt <= x, year_trt >= 1) %>%
          pull(bc.dist) %>%
          mean()
      }))
    return(output)
  }))

df.sta.max.mean.bc.numeric <- df.max.mean.bc.numeric %>%
  select(-data, -control.plot, -control.data) %>%
  unnest(cols = results)

df.sta.extent.compositional.change <- df.sta.max.mean.bc.numeric %>%
  select(-max.bc.numeric) %>%
  rename(extent.compositional.change = mean.bc.numeric)
df.sta.extent.compositional.change <- na.omit(df.sta.extent.compositional.change)
```

# Section 3: Compile stability components 

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE}
df.tidy.stabilities <- df.sta.temporal.functional.variability.mass %>%
  full_join(df.sta.temporal.compositional.variability) %>%
  full_join(df.sta.extent.functional.change.mass) %>%
  full_join(df.sta.extent.compositional.change) 
#only keep the core columns
df.tidy.stabilities <- df.tidy.stabilities %>% 
  select(site_code, block, trt, plot, end_year, contains("variability"),contains("change"))
```

# Section 4: A preliminary view of the stability data

**4.1 The frequency distribution of stability components**

First have a look at the frequency distribution of the stability components.

```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, prompt=TRUE, fig.width = 9, fig.height = 9, catch=TRUE}
logit.f <- function(x) log(x / (1 - x))

ggplot.freq.f <- function(dat, xlab = "x") {
  dat %>%
    set_colnames(value = "x") %>%
    ggplot(aes(x = x)) +
    geom_histogram(aes(y = ..density..),
      color = "black", fill = "gray90", size = 1, bins = 50
    ) +
    geom_density(color = "blue", size = 1) +
    theme_classic() +
    ylab("Density") +
    theme(axis.text = element_text(color = "black")) +
    xlab(xlab)
}

p1.1.a <- df.tidy.stabilities %>%
  select(temporal.functional.variability) %>%
  ggplot.freq.f(., xlab = "Functional variability")
p1.1.b <- df.tidy.stabilities %>%
  select(temporal.functional.variability.detrended) %>%
  ggplot.freq.f(., xlab = "Functional variability - detrended")
p1.2 <- df.tidy.stabilities %>%
  select(temporal.compositional.variability) %>%
  ggplot.freq.f(., xlab = "Compositional variability")
p1.3 <- df.tidy.stabilities %>%
  select(extent.functional.change) %>%
  ggplot.freq.f(., xlab = "Functional change")
p1.4 <- df.tidy.stabilities %>%
  select(extent.compositional.change) %>%
  ggplot.freq.f(., xlab = "Compositional change")

p2.1.a <- df.tidy.stabilities %>%
  select(temporal.functional.variability) %>%
  log() %>%
  ggplot.freq.f(., xlab = "log(functional variability)")
p2.1.b <- df.tidy.stabilities %>%
  select(temporal.functional.variability.detrended) %>%
  log() %>%
  ggplot.freq.f(., xlab = "log(functional variability - detrended)")
p2.2 <- df.tidy.stabilities %>%
  select(temporal.compositional.variability) %>%
  log() %>%
  ggplot.freq.f(., xlab = "log(compositional variability)")
p2.3 <- df.tidy.stabilities %>%
  select(extent.functional.change) %>%
  log() %>%
  ggplot.freq.f(., xlab = "log(functional change)")
p2.4 <- df.tidy.stabilities %>%
  select(extent.compositional.change) %>%
  logit.f() %>%
  ggplot.freq.f(., xlab = "logit(compositional change)")

library(patchwork)
p.distribution <- (p1.1.a + p2.1.a) / (p1.1.b + p2.1.b) / (p1.2 + p2.2) / (p1.3 + p2.3) / (p1.4 + p2.4) 
p.distribution <- p.distribution +
  plot_annotation(tag_levels = "a")
p.distribution
pdf(paste0(getwd(), "/myfigure/freqency.distribution.of.stability.components.pdf"), width = 9, height = 9)
p.distribution
dev.off()
```

Based on the distribution shape above, I transformed the stability components.
```{r, warning=FALSE, message=FALSE, eval=TRUE, highlight=T, prompt=TRUE, fig.width = 5, fig.height = 5, catch=TRUE}
df.tidy.stabilities <- df.tidy.stabilities %>% 
  mutate(func_var = log(temporal.functional.variability),
         func_var_det = log(temporal.functional.variability.detrended),
         comp_var = log(temporal.compositional.variability),
         func_change =  log(extent.functional.change),
         comp_change = logit.f(extent.compositional.change)) %>% 
  select(-contains("variability"), -contains("nal.change"))
save(df.tidy.stabilities,
  file = paste0(data.path, "/df.tidy.stabilities.Rdata")
)
```

**4.2 The correlation between two measure of temporal functional variability**

```{r, warning=FALSE, message=FALSE, eval=T, highlight=T, prompt=TRUE, fig.width = 5, fig.height = 5, catch=TRUE}
pdf(paste0(getwd(), "/myfigure/correlation.two.temporal.functional.variability.mass.pdf"), width = 5, height = 5)
# Fit a linear model
model <- lm(func_var_det ~ func_var, data = df.tidy.stabilities)

# Extract R-squared and p-value
summary_model <- summary(model)
r_squared <- summary_model$r.squared
p_value <- summary_model$coefficients[2, 4]

# Create the plot
plot(
  df.tidy.stabilities$func_var, 
  df.tidy.stabilities$func_var_det, 
  pch = 1, col = "blue",
  xlab = "Functional Variability", 
  ylab = "Detrended Functional Variability", 
)

# Add regression line
abline(model, col = "red", lwd = 2)

# Add R-squared and p-value in the upper left corner
text(
  x = -1, 
  y = 2.5,
  labels = paste0("R² = ", round(r_squared, 3), "\n", "P = ", format.pval(p_value, digits = 3)),
  adj = c(0, 1),  # Align text to top-left corner
  cex = 1  # Slightly increase text size
)

dev.off()

plot(
  df.tidy.stabilities$func_var, 
  df.tidy.stabilities$func_var_det, 
  pch = 1, col = "blue",
  xlab = "Functional Variability", 
  ylab = "Detrended Functional Variability", 
)

# Add regression line
abline(model, col = "red", lwd = 2)

# Add R-squared and p-value in the upper left corner
text(
  x = -1, 
  y = 2.5,
  labels = paste0("R² = ", round(r_squared, 3), "\n", "P = ", format.pval(p_value, digits = 3)),
  adj = c(0, 1),  # Align text to top-left corner
  cex = 1  # Slightly increase text size
)
```